---
title: "Zooplankton of Lake Champlain"
author: "Alex Naccarato"
date: "8/6/2019"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(plyr)
library(dplyr)
library(gtable) # to get two plots aligned ggplot
library(grid) # low-level grid functions are required
library(plotly)
library(reshape2)
library(lubridate)
library(plotly)

getpath4data <- function() {
  if(Sys.getenv("USER")=="Rosalie") return("/Volumes/-/Script R/LCM_GitHub_Data_LCM/")
  if(Sys.getenv("USER")=="alexnaccarato") return("~/Desktop/Food-Web 2018-2019/LCM_GitHub_Data/")
  if(Sys.getenv("USER")!="Rosalie"|Sys.getenv("USER")!="alexnaccarato") stop("You need to get the data.")
}

plktn_pre2010  <- read.delim(paste0(getpath4data(),"LCM_bio_PeteStangel/Plankton data pre 2010.txt"))
plktn_post2010 <- read.delim(paste0(getpath4data(),"LCM_bio_PeteStangel/Plankton data after 2010.txt"))
sites_raw      <- read.delim(paste0(getpath4data(),"LCM_bio_PeteStangel/Plankton data stations.txt"))
sites          <- sites_raw[sites_raw$LocationID %in% as.factor(plktn_pre2010$LocationID),]

plktn <- rbind(plktn_pre2010,plktn_post2010)

plktn$VisitDate <- parse_date_time(x = plktn$PlanktonData.VisitDate, orders = c("d-b-y","d/m/Y"))
head(plktn)

names(plktn)
plktn$PlanktonType[plktn$PlanktonType=="phyto"] <- "Phyto"

plktn$Type_SpeciesID <- paste(plktn$PlanktonType, plktn$SpeciesID, sep="_")
plktn2 <- dcast(plktn,  VisitDate + LocationID ~ Type_SpeciesID,value.var = "Result",fun.aggregate = sum, na.rm = TRUE )
head( plktn2)
dim(plktn2)
names(plktn2)

phyto <- plktn2[,c(1, 2, grep("Phyto_", names(plktn2)))]
zoo   <- plktn2[,c(1, 2, grep("Zoo_",   names(plktn2)))]

zoop_param <- read.delim(paste0(getpath4data(), "zoop_parameters.txt"))

whatsleft <- 
  names(zoo[,-c(grep("aphni",     names(zoo)),
              grep("osmin",       names(zoo)),
              grep("epto",        names(zoo)),
              grep("otifer|Anuraeopsis|Ascomorpha|Asplanchna|Brachi|Collotheca|Conochilus|Cupelopagis|Euchlanis|Filinia|Kellicottia|Keratella|Lecane|Monostyla|Nothalca|Ploesoma|Polyarthra|Synchaeta|Trichocerca",    names(zoo)),
              grep("iaphanoso",               names(zoo)),
              grep("yclop",                   names(zoo)),
              grep("alanoid|Limnocalanus",    names(zoo)),
              grep("opepo|Epischura|Harpacticoid",names(zoo)))])
zoo$BOSMINA      <- rowSums(zoo[,grep("osmin",names(zoo))])     # small grazer
zoo$DAPHNIA      <- rowSums(zoo[,grep("aphni",names(zoo))])     # medium grazer
zoo$DIAPHANOSOMA <- rowSums(zoo[,grep("iaphanoso",names(zoo))]) # medium grazer
zoo$LEPTO        <- rowSums(zoo[,grep("epto", names(zoo))])      # predator
zoo$CYCLOP       <- rowSums(zoo[,grep("yclop",names(zoo))])     # copepod
zoo$CALANOID     <- rowSums(zoo[,grep("alanoid|Limnocalanus",
                                              names(zoo))])   # copepod
zoo$ROTIFER      <- rowSums(zoo[,grep("otifer|Anuraeopsis|Ascomorpha|Asplanchna|Brachi|Collotheca|Conochilus|Cupelopagis|Euchlanis|Filinia|Kellicottia|Keratella|Lecane|Monostyla|Nothalca|Ploesoma|Polyarthra|Synchaeta|Trichocerca",names(zoo))])    # small grazer
zoo$COPEPOD      <- rowSums(zoo[,grep("opepo|Epischura|Harpacticoid",names(zoo))])    # large grazer


# Group by very large groups (i.e. grazers, predators)
zoo$PREDATOR    <- rowSums(zoo[,grep("Polyphemus|LEPTO|Holopedium",names(zoo))])   
zoo$GRAZER_sm   <- rowSums(zoo[,grep("BOSMINA|Alon|Chyd|ROTIF|leurox|ladocer",names(zoo))])  
zoo$GRAZER_lr   <- rowSums(zoo[,grep("DAPHNIA|Sida|DIAPHA|CYCLOP|CALANO|opep|Eurycercus|aupl",names(zoo))])  
zoo$PARASITIC   <- zoo[,grep("Ergasilus",names(zoo))]
  
whatsleft[-c(grep("Polyphemus|LEPTO|Holopedium",whatsleft),
            grep("BOSMINA|Sida|Alon|Chyd|ROTIF|leurox|ladocer",whatsleft),
            grep("DAPHNIA|DIAPHA|CYCLOP|CALANO|COPEPOD|Eurycercus|aupl",whatsleft),
            grep("Ergasilus",whatsleft))]

zoo$LocationID2  <- sites[match(zoo$LocationID, sites$LocationID), "StationID"]
zoo$Year         <- as.numeric(format(as.Date(zoo$VisitDate, "%Y-%m-%d"),"%Y"))
zoo$Month        <- as.numeric(format(as.Date(zoo$VisitDate, "%Y-%m-%d"),"%m"))
zoo$yday         <- as.numeric(format(as.Date(zoo$VisitDate, "%Y-%m-%d"),"%j"))

# Biomass values      <- old zoop counts           ug/m3    g/m3      g/km3        t/km3
zoo$BOSMINA_biom      <- round(zoo$BOSMINA      *  12.252 / 1000000 * 1000000000 / 1000000 , 2)
zoo$DAPHNIA_biom      <- round(zoo$DAPHNIA      * 195.881 / 1000000 * 1000000000 / 1000000 , 2)
zoo$DIAPHANOSOMA_biom <- round(zoo$DIAPHANOSOMA *  26.219 / 1000000 * 1000000000 / 1000000 , 2)
zoo$LEPTO_biom        <- round(zoo$LEPTO        * 163.714 / 1000000 * 1000000000 / 1000000 , 2)
zoo$CYCLOP_biom       <- round(zoo$CYCLOP       *  96.724 / 1000000 * 1000000000 / 1000000 , 2)
zoo$CALANOID_biom     <- round(zoo$CALANOID     * 229.355 / 1000000 * 1000000000 / 1000000 , 2)
zoo$ROTIFER_biom      <- round(zoo$ROTIFER      *   7.861 / 1000000 * 1000000000 / 1000000 , 2)
zoo$COPEPOD_biom      <- round(zoo$COPEPOD      *  19.126 / 1000000 * 1000000000 / 1000000 , 2)
zoo$PREDATOR_biom     <- round(zoo$PREDATOR     * 213.582 / 1000000 * 1000000000 / 1000000 , 2)
zoo$GRAZER_sm_biom    <- round(zoo$GRAZER_sm    *  21.039 / 1000000 * 1000000000 / 1000000 , 2)
zoo$GRAZER_lr_biom    <- round(zoo$GRAZER_lr    * 187.096 / 1000000 * 1000000000 / 1000000 , 2)
zoo$PARASITIC_biom    <- round(zoo$PARASITIC    *  35.039 / 1000000 * 1000000000 / 1000000 , 2)
```

## Mean wet weight of each large group

```{r}
for (i in unique(zoop_param$Large_group)) {
  print(paste0(i, ": ", round(mean(zoop_param$WW__g[zoop_param$Large_group == i], na.rm = TRUE), 3)))
}
```

## Mean wet weight of each small group

```{r}
for (i in unique(zoop_param$Small_group)) {
  print(paste0(i, ": ", (round(mean(zoop_param$WW__g[zoop_param$Small_group == i], na.rm = TRUE), 3))))
}
```

## Distribution of the average size of zooplankton

Here, we're plotting two very similar plots. The first contains the zooplankton <i> Eurycercus spp. </i> which has an average wet weight of over 3,000 micrograms. The second does not contain this same data point so the viewer can better observe the majority of the data.

```{r}
ggplot(data = zoop_param, mapping = aes(x = WW__g)) +
  geom_histogram(bins = 30) +
  labs(x = expression(Wet~Weight~(mu*'g')), y = "Count")

zoop_param <- zoop_param[zoop_param$WW__g < 2000,]

ggplot(data = zoop_param, mapping = aes(x = WW__g)) +
  geom_histogram(bins = 30) +
  labs(x = expression(Wet~Weight~(mu*'g')), y = "Count")

zoop_param <- read.delim(paste0(getpath4data(), "zoop_parameters.txt"))
```

## Distribution of the abundance and biomass of zooplankton 

The first plot displays the total zooplankton count over time. The second plot displays the total zooplankton biomass over time. 

```{r}
for (i in unique(zoo$Year)) {
  print(paste0(i, ": ", sum(rowSums(zoo[zoo$Year == i, 84:91])), " zooplankton"))
}

ggplot(data = zoo, mapping = aes(x = zoo$Year, y = rowSums(zoo[,84:91]))) +
  geom_bar(stat = "identity") +
  labs(y = "Total Zooplankton Count")

for (i in unique(zoo$Year)) {
  print(paste0(i, ": ", sum(rowSums(zoo[zoo$Year == i, 100:107])), " t/km3"))
}

ggplot(data = zoo, mapping = aes(x = zoo$Year, y = rowSums(zoo[, 100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Year")
```

## Intrannual changes in zooplankton biomass

These plots show the intraannual changes, or the changes within each year, in zooplankton biomass. 

```{r}
# Can't currently figure out how to only choose one year for this plot:

ggplot(data = zoo, mapping = aes(x = zoo$Month, y = rowSums(zoo[, 100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month")

zoo_1992 <- as.data.frame(zoo[zoo$Year == 1992,])
zoo_1993 <- as.data.frame(zoo[zoo$Year == 1993,])
zoo_1994 <- as.data.frame(zoo[zoo$Year == 1994,])
zoo_1995 <- as.data.frame(zoo[zoo$Year == 1995,])
zoo_1996 <- as.data.frame(zoo[zoo$Year == 1996,])
zoo_1997 <- as.data.frame(zoo[zoo$Year == 1997,])
zoo_1998 <- as.data.frame(zoo[zoo$Year == 1998,])
zoo_1999 <- as.data.frame(zoo[zoo$Year == 1999,])
zoo_2000 <- as.data.frame(zoo[zoo$Year == 2000,])
zoo_2001 <- as.data.frame(zoo[zoo$Year == 2001,])
zoo_2002 <- as.data.frame(zoo[zoo$Year == 2002,])
zoo_2003 <- as.data.frame(zoo[zoo$Year == 2003,])
zoo_2004 <- as.data.frame(zoo[zoo$Year == 2004,])
zoo_2005 <- as.data.frame(zoo[zoo$Year == 2005,])
zoo_2006 <- as.data.frame(zoo[zoo$Year == 2006,])
zoo_2007 <- as.data.frame(zoo[zoo$Year == 2007,])
zoo_2008 <- as.data.frame(zoo[zoo$Year == 2008,])
zoo_2009 <- as.data.frame(zoo[zoo$Year == 2009,])
zoo_2010 <- as.data.frame(zoo[zoo$Year == 2010,])
zoo_2011 <- as.data.frame(zoo[zoo$Year == 2011,])
zoo_2012 <- as.data.frame(zoo[zoo$Year == 2012,])
zoo_2013 <- as.data.frame(zoo[zoo$Year == 2013,])
zoo_2014 <- as.data.frame(zoo[zoo$Year == 2014,])
zoo_2015 <- as.data.frame(zoo[zoo$Year == 2015,])

ggplot(data = zoo_1993, mapping = aes(x = zoo_1993$Month, y = rowSums(zoo_1993[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 1993)

ggplot(data = zoo_1994, mapping = aes(x = zoo_1994$Month, y = rowSums(zoo_1994[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 1994)

ggplot(data = zoo_1995, mapping = aes(x = zoo_1995$Month, y = rowSums(zoo_1995[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 1995)

ggplot(data = zoo_1996, mapping = aes(x = zoo_1996$Month, y = rowSums(zoo_1996[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 1996)

ggplot(data = zoo_1997, mapping = aes(x = zoo_1997$Month, y = rowSums(zoo_1997[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 1997)

ggplot(data = zoo_1998, mapping = aes(x = zoo_1998$Month, y = rowSums(zoo_1998[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 1998)

ggplot(data = zoo_1999, mapping = aes(x = zoo_1999$Month, y = rowSums(zoo_1999[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 1999)

ggplot(data = zoo_2000, mapping = aes(x = zoo_2000$Month, y = rowSums(zoo_2000[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2000)

ggplot(data = zoo_2001, mapping = aes(x = zoo_2001$Month, y = rowSums(zoo_2001[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2001)

ggplot(data = zoo_2002, mapping = aes(x = zoo_2002$Month, y = rowSums(zoo_2002[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2002)

ggplot(data = zoo_2003, mapping = aes(x = zoo_2003$Month, y = rowSums(zoo_2003[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2003)

ggplot(data = zoo_2004, mapping = aes(x = zoo_2004$Month, y = rowSums(zoo_2004[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2004)

ggplot(data = zoo_2005, mapping = aes(x = zoo_2005$Month, y = rowSums(zoo_2005[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2005)

ggplot(data = zoo_2006, mapping = aes(x = zoo_2006$Month, y = rowSums(zoo_2006[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2006)

ggplot(data = zoo_2007, mapping = aes(x = zoo_2007$Month, y = rowSums(zoo_2007[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2007)

ggplot(data = zoo_2008, mapping = aes(x = zoo_2008$Month, y = rowSums(zoo_2008[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2008)

ggplot(data = zoo_2009, mapping = aes(x = zoo_2009$Month, y = rowSums(zoo_2009[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2009)

ggplot(data = zoo_2010, mapping = aes(x = zoo_2010$Month, y = rowSums(zoo_2010[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2010)

ggplot(data = zoo_2011, mapping = aes(x = zoo_2011$Month, y = rowSums(zoo_2011[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2011)

ggplot(data = zoo_2012, mapping = aes(x = zoo_2012$Month, y = rowSums(zoo_2012[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2012)

ggplot(data = zoo_2013, mapping = aes(x = zoo_2013$Month, y = rowSums(zoo_2013[,100:107]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month", title = 2013)

```

