---
title: "Lake Trout Diet Analysis"
author: "Alex Naccarato"
date: "5/14/2019"
output: html_document
---

## The Basics

I always have to include the code for my libraries so everything functions properly. In addition, I'm reading in the data file I'll be using for the analysis. I'm also getting rid of two outliers. They were very long fish that were quite skinny. 

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(binr)
library(matrixStats)

LTdata <- read.csv("/Users/alexnaccarato/Desktop/Food-Web 2018-2019/LCM_GitHub_Data/data_from_Pascal/LT_2016-2018.csv")

LTdata$Weight[LTdata$Weight < 3000 & LTdata$Total.Length > 750] <- NA
LTdata$Total.Length[LTdata$Weight < 3000 & LTdata$Total.Length > 750] <- NA
```

## Relationship Between Length and Weight, Separated by Age

Here, I'm creating a scatterplot of the weight of the fish by their length. I included a filter for age to see how this relationship varies by age. In addition, I included a log-transformed version of this plot. 

```{r age, warning=FALSE}
LTdata %>% ggplot(mapping = aes(x = Total.Length,
                                y = Weight, 
                                color = Age)) +
  geom_point() +
  labs(x = "Total Length (mm)", y = "Weight (g)") +
  geom_smooth(method = "loess", color = "black")

LTdata %>%  ggplot(mapping = aes(x = Total.Length, 
                                 y = Weight,
                                 color = Age)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Total Length (mm)", y = "Weight (g)") +
  geom_smooth(method = "lm", color = "black")
```

## Relationship Between Length and Weight, Separated by Clip Location

Here, I'm creating another scatterplot of the weight/length relationship for lake trout, but included the location of the clip as a filter. I also included a log-transformed version of this plot.

```{r cliplocation, warning=FALSE}
LTdata %>% ggplot(mapping = aes(x = Total.Length,
                                y = Weight, 
                                color = Clip.Location)) +
  geom_point() +
  labs(x = "Total Length (mm)", y = "Weight (g)", color = "Clip Location") +
  geom_smooth(method = "loess", color = "black")

LTdata %>%  ggplot(mapping = aes(x = Total.Length, 
                                 y = Weight,
                                 color = Clip.Location)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Total Length (mm)", y = "Weight (g)", color = "Clip Location") +
  geom_smooth(method = "lm", color = "black")
```

## Length and Weight Relationship between Stocked and Wild Lake Trout

Stocked trout are typically larger than wild fish. They often have longer lengths and greater weights than wild fish. 

```{r stocked, warning = FALSE}
LTdata %>% mutate(group = case_when(Clip.Location == "NC" ~ "Wild",
                                    Clip.Location == "AD" ~ "Stocked",
                                    Clip.Location == "LP" ~ "Stocked",
                                    Clip.Location == "LV" ~ "Stocked",
                                    Clip.Location == "LV*" ~ "Stocked",
                                    Clip.Location == "LVRV" ~ "Stocked",
                                    Clip.Location == "RP" ~ "Stocked",
                                    Clip.Location == "RV" ~ "Stocked")) %>% 
  ggplot(mapping = aes(x = Total.Length,
                                y = Weight, 
                                color = group)) +
  geom_point() +
  labs(x = "Total Length (mm)", y = "Weight (g)", color = "Clip Location") +
  geom_smooth(method = "loess", color = "black")

LTdata %>% mutate(group = case_when(Clip.Location == "NC" ~ "Wild",
                                    Clip.Location == "AD" ~ "Stocked",
                                    Clip.Location == "LP" ~ "Stocked",
                                    Clip.Location == "LV" ~ "Stocked",
                                    Clip.Location == "LV*" ~ "Stocked",
                                    Clip.Location == "LVRV" ~ "Stocked",
                                    Clip.Location == "RP" ~ "Stocked",
                                    Clip.Location == "RV" ~ "Stocked")) %>% 
  ggplot(mapping = aes(x = Total.Length,
                       y = Weight, 
                       color = group)) +
  geom_point() +
  labs(x = "Total Length (mm)", y = "Weight (g)", color = "Clip Location") +
  geom_smooth(method = "loess", color = "black") +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Total Length (mm)", y = "Weight (g)", color = "Clip Location") +
  geom_smooth(method = "lm", color = "black")
```

## Proportion of Empty Stomachs For Each Length Class

Here, I'm creating a bar chart of the proportion of empty stomachs size classes of lake trout. It appears that the likelihood of finding a trout with an empty stomach is more likely in older fish. 

```{r empty.by.lengthclass}
LTdata <- LTdata  %>% mutate(size_class = case_when(
  Total.Length <= 100 ~ "[0,100]",
  Total.Length > 100 & Total.Length <= 200 ~ "]100,200]", 
  Total.Length > 200 & Total.Length <= 300 ~ "]200,300]",
  Total.Length > 300 & Total.Length <= 400 ~ "]300,400]",
  Total.Length > 400 & Total.Length <= 500 ~ "]400,500]",
  Total.Length > 500 ~ "]500,∞]"))

(summ_sc <- with(LTdata, tapply(rep(1,nrow(LTdata)),list("Size class" = size_class, "Food" = Food.in.Stomach), sum)))

(summ_sc <- as.data.frame(summ_sc / rowSums(summ_sc, na.rm=T) * 100))

ggplot(data = summ_sc, aes(x = rownames(summ_sc),y = summ_sc[,4])) +
  geom_bar(stat = "identity") + xlab("Size Class") + ylab("Percentage of Empty Stomachs")
```

## Lake Trout Diet by Length

The relationship between the size of stomachs and the length of the fish is not well summarized by a linear model. 

```{r stomach.weight.by.length, warning=FALSE}
LTdata %>% ggplot(mapping = aes(x = Total.Length, y = as.numeric(Stomach.Weight))) +
  geom_point() + 
  xlab("Length (mm)") + 
  ylab("Stomach Weight (g)") +
  stat_smooth(method = "lm") 
```

## Lake Trout Diet by Weight

The relationship between the size of the stomachs and the weight of the fish is not well summarized by a linear model. 

```{r stomach.weight.by.weight, warning=FALSE}
LTdata %>% ggplot(mapping = aes(x = Weight, y = as.numeric(Stomach.Weight))) +
  geom_point() + 
  xlab("Length (mm)") + 
  ylab("Stomach Weight (g)") +
  stat_smooth(method = "lm") 
```


