---
title: "Model parameter EwE Lake Champlain"
author: "Rosalie Bruel"
date: "11/07/2019"
output: 
  html_document: 
    df_print: paged
    fig_caption: yes
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
  df_print: paged
fontsize: 11pt
---

_Last update: `r Sys.Date()`_

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The objective of this document is to gather all in one place the decisions made for the parameters of the model.
I'm starting this document on July 16th, 2019, having entered some parameters over the past 4 months and realizing I already forgot where I got them from... Hopefully I'll find my notes back, otherwise, having everything in one place might be easier.

# Detritus

We use Pauly et al (1993) equation to calculate the biomass of detritus:  
$Log10(D) = 0.954 * log10(PP) + 0.863 * log10(E) – 2.41$,

with E= depth of the euphotic zone.
I'm calculating the depth of the euphotic zone based on the secchi depth.

## Secchi Depth

The secchi depth across all sites and years

```{r echo=FALSE}
secchi_depth_p <- total %>% ggplot(mapping = aes(x = as.factor(StationID), y = `Secchi Depth`, fill = StationID)) +
  geom_boxplot() +
  xlab("Station ID") +
  ylab("Secchi depth (m)") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle(paste0("Secchi depths across all sites and years"))
secchi_depth_p

eu_z = 4.9
```
We set the euphotic zone for the main lake (StationID= 16 & 19) at `r eu_z` m.



```{r}
D = 10^(0.954 * log10(1) + 0.863 * log10(eu_z) - 2.41)
D
```

