---
title: "GIS exploration"
author: "Rosalie Bruel"
date: "2020-01-22 -- last updated on `r Sys.Date()` "
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    df_print: paged
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
fontsize: 11pt
editor_options:
  chunk_output_type: console
  df_print: paged
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = dirname(getwd()))
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plotly)
library(plyr) #for ddply
library(dplyr)
library(lubridate) # for date time
library(changepoint) #changepoint analysis
library(reshape2) # for melt
library(Hmisc) # for err bar plot
library(FSA);library(FSAdata) # Fisheries stock assessment methods and data
library(leaflet)
library(janitor)
library(rgdal)
library(raster)
library(elevatr) # for elevation USA
library(sf)
library(USAboundaries) # counties and states delimitation USA
library(rnaturalearth)
library(rasterVis) # for levelplot
library(GISTools) # for North arrow in plot (function north.arrow())
library(RColorBrewer)
library(grid)
library(gridExtra) # get two plots side by side
library(knitr) # for kable
library(stringr) # for str_replace()
library(wesanderson) # colors
library(ggraph) # for nodes plot
library(igraph) # probably better for nodes graph because can also assign weigth
library(tidygraph) # to transform dataframe in node datasets
library(scales) # Get full number in x/y scales with scale_x_continuous(labels = comma)/scale_y_continuous(labels = comma) instead of 1.00e09 for example
library(concurve) # plot p-value and s-value, see recommandations in Chow and Sander, 2019
library(psych) # for pairs-panels (scaterplot matrix)

# Function to draw North arrow
#https://gis.stackexchange.com/questions/243743/alternative-north-arrow-using-r-gistools-package
northarrow <- function(loc,size,bearing=0,cols,cex=1,...) {
  # checking arguments
  if(missing(loc)) stop("loc is missing")
  if(missing(size)) stop("size is missing")
  # default colors are white and black
  if(missing(cols)) cols <- rep(c("white","black"),8)
  # calculating coordinates of polygons
  radii <- rep(size/c(1,4,2,4),4)
  x <- radii[(0:15)+1]*cos((0:15)*pi/8+bearing)+loc[1]
  y <- radii[(0:15)+1]*sin((0:15)*pi/8+bearing)+loc[2]
  # drawing polygons
  for (i in 1:15) {
    x1 <- c(x[i],x[i+1],loc[1])
    y1 <- c(y[i],y[i+1],loc[2])
    polygon(x1,y1,col=cols[i])
  }
  # drawing the last polygon
  polygon(c(x[16],x[1],loc[1]),c(y[16],y[1],loc[2]),col=cols[16])
  # drawing letters
  b <- c("","N","","")
  #b <- c("E","N","W","S")
  for (i in 0:3) text((size+par("cxy")[1])*cos(bearing+i*pi/2)+loc[1],
                      (size+par("cxy")[2])*sin(bearing+i*pi/2)+loc[2],b[i+1],
                      cex=cex)
}

# Captions with library captioner
# https://community.rstudio.com/t/avoiding-repetitive-latex-codes-in-r-markdown/7834/12
# https://cran.r-project.org/web/packages/captioner/vignettes/using_captioner.html 
fig_cap <- captioner::captioner()
tab_cap <- captioner::captioner("Table")

getpath4data <- function() {
  if(Sys.getenv("USER")=="Rosalie") return("/Volumes/-/Script R/LCM_GitHub_Data_LCM/")
  if(Sys.getenv("USER")=="alexnaccarato") return("~/Desktop/Food-Web 2018-2019/LCM_GitHub_Data/")
    if(Sys.getenv("USER")=="YOUR USER NAME") return("YOUR PATH")
  if(Sys.getenv("USER")!="Rosalie"|Sys.getenv("USER")!="alexnaccarato") stop("You need to get the data.")
}

# turn this to true so some stuffs are not run when I'm knitting (and vice versa)
R_U_KNITTING = TRUE 
```


Using tuto <a href="https://translate.google.com/translate?hl=fr&sl=en&u=https://www.neonscience.org/raster-data-r&prev=search">here</a>.

```{r message=FALSE, warning=FALSE}
DEM <- raster(paste0(getpath4data(),"GIS/LCbathy.tif"))
# calculate and save the min and max values of the raster to the raster object
DEM <- setMinMax(DEM)

#Get min and max cell values from raster #NOTE: this code may fail if the raster is too large
# file is too large, but below would be the function cellStats
#cellStats(DEM, range)

#view coordinate reference system
DEM@crs
#nothing. Use coordinates from ne_countries, package sf
world <- ne_countries(scale="large", returnclass = "sf")
DEM <- projectRaster(DEM, crs = crs(world))
DEM@crs

# view raster extent
DEM@extent

# the distribution of values in the raster - making sure there are no really weird values
hist(DEM, main="Distribution of elevation values", col= "purple")

# plot the raster
plot(DEM, main="Digital Elevation Model, Lake Champlain")

# create a plot of our raster using function image() that works better for large object (because plot has a default of 100,000 pixels)
# specify the range of values that you want to plot in the DEM # just plot pixels between 250 and 300 m in elevation
image(DEM, zlim=c(-80,0))

# we can specify the colors too
col <- terrain.colors(5)
par(new=T)
image(DEM, zlim=c(-10,0), main="Digital Elevation Model (DEM)", col=col)

# add a color map with 5 colors
col=terrain.colors(5)
col=colorspace::diverge_hcl(5)

# add breaks to the colormap (6 breaks = 5 segments) 
brk <- c(-120, -80, -60, -40, -20, 0) 
plot(DEM, col=col, breaks=brk, main="DEM with more breaks")

#edit legend appearance
# First, expand right side of clipping rectangle to make room for the legend
# turn xpd off
par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 4.5)) 
# Second, plot w/ no legend 
plot(DEM, col=col, breaks=brk, main="DEM with a Custom (but flipped) Legend", legend = FALSE) 
# Third, turn xpd back on to force the legend to fit next to the plot. 
par(xpd = TRUE) 
# Fourth, add a legend - & make it appear outside of the plot 
legend( par()$usr[2], 45, legend = c("Shallowest", "Shallower yet", "Middle","A bit deeper", "Deepest"), fill = rev(col))
# Fifth, turn xpd = False)
par(xpd = FALSE) 


# Crop a figure
#define the crop extent
cropbox2 <-c(-73.35,-73.2,44.4,44.5)
#crop the raster
DEMcrop2 <- crop(DEM, cropbox2) 
#plot cropped DEM
plot(DEMcrop2)
```

# Smelt paper map 

```{r smelt map}
# Read limit LC basin
lcb = st_read(paste0(getpath4data(),"GIS/LCB_boundary_2013/LCB_boundary_2013.shp"))
lcb <- st_transform(lcb, crs = crs(world))
plot(lcb, axes=T, col=adjustcolor("grey", alpha.f = .3), main="Lake Champlain Basin")
lcb_df <- as.data.frame(lcb$geometry) 
lcb_df <- unlist(lcb_df)
lcb_df <- data.frame("lon" = lcb_df[1:(length(lcb_df)/2)], 
                     "lat" = lcb_df[(length(lcb_df)/2+1):length(lcb_df)])
plot(lcb_df, type="l")
par(new=T)
plot(DEM, col=col, main="")#, axes="F", add = TRUE) 
lines(lcb_df)


brk <- c(-122,-100, -80, -60, -50, -40, -30, -20, -10,0) 
#brk <- c(-120, -80, -60, -40, -20, 0) 

col=colorspace::diverge_hcl(c(length(brk)-1), l = 60)
col=colorspace::diverge_hsv(c(length(brk)-1), s = .8)
colr <- colorRampPalette(brewer.pal(11, 'RdYlBu'))
col=rev(colr(length(brk)))
#plot(1:10,1:10, pch=16, cex=4, col=col)

#pdf(paste0(getwd(),"/Output/Figures/9-Smelt/Smelt_sampling_sites.pdf"),width = 6, height = 9)
if(!R_U_KNITTING) png(paste0(getwd(),"/Output/Figures/9-Smelt/Smelt_sampling_sites.png"), width = 5.5, height = 7.5, units = "in", res = 120)
par(mar=c(5.1,5.1,2.1,2.1))
plot(DEM, col=col, main="", axes="F") 
axis(1, at =  seq(-74.5,-72, by=.5), labels = paste0(abs(seq(-74.5,-72, by=.5)), "°W"))
axis(2, at =  seq(43,45, by=.5), labels = paste0(seq(43,45, by=.5), "°N"), las=1)
mtext("Longitude", side = 1, line = 2.5)
mtext("Latitude", side = 2, line = 3.6)
#levelplot(DEMcrop_Ben, par.settings = BuRdTheme, contour=TRUE)
# plot (add smelt sampling coordinates on top of it)
smelt_coord <- read.csv(file = paste0(getpath4data(), "smelt_coord.csv"))
smelt_coord$Basin <- c("Main", "NEArm", "Main", "Malletts", rep("Main", 6), "NEArm")
smelt_coord <- smelt_coord[order(smelt_coord$Latitude, decreasing = T),]
points(smelt_coord$Longitude[smelt_coord$Basin=="Main"],
       smelt_coord$Latitude[smelt_coord$Basin=="Main"], pch=16, cex=1.2, col="white")
points(smelt_coord$Longitude[smelt_coord$Basin=="NEArm"],
       smelt_coord$Latitude[smelt_coord$Basin=="NEArm"], pch=17, cex=1.2, col="white")
points(smelt_coord$Longitude[smelt_coord$Basin=="Malletts"],
       smelt_coord$Latitude[smelt_coord$Basin=="Malletts"], pch=15, cex=1.2, col="white")

points(smelt_coord$Longitude[smelt_coord$Basin=="Main"],
       smelt_coord$Latitude[smelt_coord$Basin=="Main"], pch=16)
points(smelt_coord$Longitude[smelt_coord$Basin=="NEArm"],
       smelt_coord$Latitude[smelt_coord$Basin=="NEArm"], pch=17)
points(smelt_coord$Longitude[smelt_coord$Basin=="Malletts"],
       smelt_coord$Latitude[smelt_coord$Basin=="Malletts"], pch=15)
TeachingDemos::shadowtext(smelt_coord$Longitude-0.045, smelt_coord$Latitude-0.015, 1:nrow(smelt_coord), col = "black",bg = "white", theta = seq(pi/4, 2 * pi, length.out = 8), r = 0.15, cex=1.1)
legend("bottomleft", legend = c("Main Lake", "Malletts Bay", "Northeast Arm"), pch= c(16,15,17), cex=.9)
scalebar(25, xy=c(-73.18, 43.6), type='bar', divs=2, below = "kilometers")
northarrow(loc=c(-73.0, 43.78), size = .09)

if(!R_U_KNITTING) dev.off()
par(mar=c(5.1,4.1,4.1,2.1))

# for legend within paper
# paste("(",1:nrow(smelt_coord),") ", smelt_coord$Station, sep="", collapse = ", ")

```


```{r plot bathy with color, echo=FALSE, message=FALSE, warning=FALSE}
# Define a colour ramp palette
colr <- colorRampPalette(brewer.pal(11, 'RdYlBu'))

p <- levelplot(bathy2, 
          margin=TRUE,    # turn to F to suppress marginal graphics
          colorkey=list(
            space='left',  # where to plot legend
            labels=list(at=seq(-120,0,20))      # legend ticks and labels
            ),    
          par.settings=list(
            axis.line=list(col='black') # turn to transparent to suppress axes and legend outline
          ),
          scales=list(draw=TRUE),  # turn to F to suppress axis labels
          col.regions=colr,         # colour ramp
          at=seq(-130, -1, len=130))  # colour ramp breaks

# p$legend$top

p 

#pdf(paste0(getwd(),"/Output/Figures/1-descriptive/LG_bathy.pdf"),width = 4, height = 9)
pdf(paste0(getwd(),"/Output/Figures/1-descriptive/LG_bathy.pdf"),width = 9, height = 7)
p
dev.off()

# pdf(paste0(getwd(),"/Output/Figures/1-descriptive/LC_bathy_BW.pdf"),width = 4, height = 9)
# plot(bathy2,
#      main = "Bathymetric map of Lake Champlain",
#      col = grey(1:100/100),
#      axes = FALSE,
#      box = FALSE)
# dev.off()


histogram(bathy2, main='Lake Champlain depth histogram', xlab='depth (m)')

```


```{r map for Lake Geneva, eval=FALSE, include=FALSE}
colr <- colorRampPalette(brewer.pal(11, 'RdYlBu'))

pchamp <- levelplot(DEM,  margin=TRUE,colorkey=list(
            space='left',  # where to plot legend
            labels=list(at=seq(-320,0,20))      # legend ticks and labels
            ),    
          par.settings=list( axis.line=list(col='black') ),
          scales=list(draw=TRUE), col.regions=colr, at=seq(-325, -1, len=101))

pchamp 


#hist plot for LC
histogram(DEM, main='Lake Champlain depth histogram', xlab='depth (m)')

```

# Map for Ben

```{r}

# add breaks to the colormap (6 breaks = 5 segments) 
brk <- seq(-26, 0, by=2) 
shadesOfGrey <- colorRampPalette(c("grey100","grey16"))
col=shadesOfGrey(length(brk))
cropbox2 <-c(-73.33,-73.27,44.64,44.68)
#crop the raster
DEMcrop_Ben <- crop(DEM, cropbox2) 
#plot cropped DEM

levelplot(DEMcrop_Ben, par.settings = BuRdTheme, contour=TRUE)

contour_ben <- rasterToContour(DEMcrop_Ben)
str(contour_ben)
plot(DEMcrop_Ben)
plot(contour_ben, add=TRUE)
as.data.frame(contour_ben)

pdf(paste0("/Users/Rosalie/Downloads/Ben_Map_bw.pdf"), height=6, width=6.45)
plot(DEMcrop_Ben, col=col)
plot(contour_ben, add=TRUE)
scalebar(2, xy=c(-73.299, 44.643), type='bar', divs=2, label = c("0","1","2 km"))
dev.off()


```



# Zebra mussels vs quagga mussel distribution

## Lake Champlain

```{r}
col=terrain.colors(3)
brk <- c(-120, -80, -50, 0) 

if(!R_U_KNITTING) pdf(paste0(getwd(),"/Output/Figures/7-ZebraMussels/LC_ZM_QM_distribution .pdf"))
par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 15)) 
# Main plot
plot(DEM, col=rev(col), breaks=brk, main="Distribution of quagga vs. zebra mussels", legend = FALSE, xlim=c(-73.6, -73.0), ylim=c(43.43, 45.18)) 
scalebar(25, xy=c(-73.2, 43.6), type='bar', divs=2)
north.arrow(-72.9, 43.77,len = .015, lab = "N", col="black")
# Add legend
legend(par()$usr[2], 45, bty = "n", 
   legend=c("Potential distribution of\n zebra mussels (<50m)","Potential distribution of\n quagga mussels (<80m)", "Too deep for both species"), 
   fill=col,
   xpd = TRUE, ncol = 1, y.intersp=2)

if(!R_U_KNITTING) dev.off()

```

## Comparison with Lake Michigan

Bunnel et al 2018 report
```{r}
michigan <- raster("/Users/Rosalie/Desktop/Travail/QGIS/1 - Data SIG/michigan_lld/michigan_lld.tif")
michigan <- projectRaster(michigan, crs = crs(world))
michigan
pmich <- levelplot(michigan,  margin=TRUE,colorkey=list(
            space='left',  # where to plot legend
            labels=list(at=seq(-320,0,20))      # legend ticks and labels
            ),    
          par.settings=list( axis.line=list(col='black') ),
          scales=list(draw=TRUE), col.regions=colr, at=seq(-325, -1, len=101))

pmich 

histogram(michigan, main='Lake Michigan depth histogram', xlab='depth (m)',xlim=c(-300,0),ylim=c(0,1.05), breaks=300)

col=terrain.colors(3)
brk <- c(-300, -80, -50, 0) 

if(!R_U_KNITTING) pdf(paste0(getwd(),"/Output/Figures/7-ZebraMussels/LMich_ZM_QM_distribution .pdf"))
par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 10)) 
# Main plot
plot(michigan, col=rev(col), breaks=brk, main="Distribution of quagga vs. zebra mussels", legend = FALSE, xlim=c(-88.5, -84), ylim=c(41.1, 46.6)) 
scalebar(100, xy=c(-86, 41.7), type='bar', divs=8)
north.arrow(-85, 42.1,len = .07, lab = "N", col="black")
# Add legend
legend(par()$usr[2], 45, bty = "n", 
   legend=c("Potential distribution of\n zebra mussels (<50m)","Potential distribution of\n quagga mussels (<80m)", "Too deep for both species"), 
   fill=col,
   xpd = TRUE, ncol = 1, y.intersp=2)

if(!R_U_KNITTING) dev.off()


str(michigan)
hist(michigan, breaks=seq(-300,350,by=1))
abline(v=0)

```


## Plot average abundance for each sampling site on map

```{r}
ZM14_stats_sites_live <- read.delim(paste0(getwd(),"/Output/Data/zebra_mussels_2014_summary_data.txt"))
ZM14_stats_sites_live$Avg_transformed <- log(ZM14_stats_sites_live$Avg)

col=terrain.colors(3)
brk <- c(-120, -80, -50, 0) 

if(!R_U_KNITTING) pdf(paste0(getwd(),"/Output/Figures/7-ZebraMussels/LC_ZM_QM_distribution_with_2014_abundance.pdf"))
par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 15)) 
# Main plot
plot(DEM, col=rev(col), breaks=brk, main="Distribution of quagga vs. zebra mussels", legend = FALSE, xlim=c(-73.6, -73.0), ylim=c(43.43, 45.18)) 
scalebar(25, xy=c(-73.2, 43.6), type='bar', divs=2)
north.arrow(-72.9, 43.77,len = .015, lab = "N", col="black")
# Add legend
legend(par()$usr[2], 45, bty = "n", 
   legend=c("Potential distribution of\n zebra mussels (<50m)","Potential distribution of\n quagga mussels (<80m)", "Too deep for both species"), 
   fill=col,
   xpd = TRUE, ncol = 1, y.intersp=2)

legend(par()$usr[2], 44.3, bty = "n",
   legend=c(max(      ZM14_stats_sites_live$Avg/2, na.rm=T),
                 mean(ZM14_stats_sites_live$Avg/2, na.rm=T),
                  min(ZM14_stats_sites_live$Avg/2, na.rm=T)), 
   pch=16, pt.cex=c(max(ZM14_stats_sites_live$Avg_transformed/2, na.rm=T),
                 mean(ZM14_stats_sites_live$Avg_transformed/2, na.rm=T),
                 min(ZM14_stats_sites_live$Avg_transformed/2, na.rm=T)),
  xpd = TRUE, ncol = 1, y.intersp=1.2)

# Add point abundance
for (i in 1:nrow(ZM14_stats_sites_live)) {
  points(ZM14_stats_sites_live$Lon[i],ZM14_stats_sites_live$Lat[i], pch=16, cex=ZM14_stats_sites_live$Avg_transformed[i]/2)
}

if(!R_U_KNITTING) dev.off()
```


# Elevation map

(not run, extraction takes a while. See files for output figure.)
```{r plot elevation map, message=FALSE, warning=FALSE, include=FALSE}
#from help here: https://gis.stackexchange.com/questions/293993/plotting-and-analyzing-extracted-elevation-data-in-r

counties <- us_counties(map_date = "1930-01-01", resolution = "high", states = c("NY","VT"))
states <- us_states(map_date = "1930-01-01", resolution = "high", states = c("NY","VT"))

counties_sf <- as(counties, "Spatial")
states_sf <- as(counties, "Spatial")
elevation_data <-get_elev_raster(states, z=9, src = "aws")
#map <- extract(elevation_data, counties)
if(!R_U_KNITTING) pdf(paste0(getwd(),"/Output/Figures/1-descriptive/map_elevation_NY_VT.pdf"), width = 6.87, height = 5)
plot(elevation_data, axes=TRUE)
#plot(st_geometry(world), add=TRUE, lwd=.7)
plot(st_geometry(states), add=TRUE)
plot(DEM, col="#9CD3DB", legend = FALSE, add = T) 
TeachingDemos::shadowtext(-73.3,44.6, labels = "Lake Champlain", col = "black",bg = "#9CD3DB", theta = seq(pi/4, 2 * pi, length.out = 8), r = 0.1, cex=.9, pos=2)
scalebar(200, xy=c(-79, 40.5), type='bar', divs=4)
north.arrow(-79.7, 40.5,len = .09, lab = "N", col="black")
if(!R_U_KNITTING) dev.off()

```

Extract depth for each location
```{r try extract depth for each location - not working rn, eval=FALSE, include=FALSE}
# From help https://gis.stackexchange.com/questions/279079/extracting-value-of-raster-with-coordinates-using-r
# xy <- xyFromCell(DEM, c(0:ncell(DEM)+1))
# xy <- data.frame(xy)
# dim(xy)
# mbreak <- seq(1, dim(xy)[1], by=100000)
# xyz <- extract(DEM, xy[1:100,], df=T)
# xyz <- cbind(xyz, xy)
# xyz<-getValues(DEM)
# write.table(xyz, file = paste0(dirname(getwd()), "/Output/Data/LakeChamplain_bathy_xyz.txt"))

xy <- xyFromCell(DEM, c(0:ncell(DEM)+1))
xyzlist <- list(x=xy[,'x'],y=xy[,'y'],z=as.matrix(r))
xyzlist <- list(x=coords[,'x'],y=coords[,'y'],z=as.matrix(r))
xyz <- as.matrix(DEM[1:5])

xyz <- raster()
xyz[] <- runif(ncell(xyz))
```


# Try getting area per depth

```{r}
DEM_hist <- hist(DEM, main="Distribution of elevation values", col= "purple", plot=F)
sum(DEM_hist$counts)
ncell(DEM)

mich_hist <-hist(michigan, main="Distribution of elevation values", col= "purple", plot=F, breaks=seq(-300,350,5))


points(DEM_hist$mids,DEM_hist$counts)

DEM_hist$percent_above <- rep(NA, length(DEM_hist$count))
for (i in 2:length(DEM_hist$breaks)) DEM_hist$percent_above[i-1] <- sum(DEM_hist$counts[DEM_hist$mids>DEM_hist$breaks[i]])/sum(DEM_hist$counts)

mich_hist$percent_above <- rep(NA, length(mich_hist$count))
for (i in 2:length(mich_hist$breaks)) mich_hist$percent_above[i-1] <- sum(mich_hist$counts[mich_hist$mids>mich_hist$breaks[i]&mich_hist$mids<0])/sum(mich_hist$counts[mich_hist$mids<0])


plot(mich_hist$mids, (1-mich_hist$percent_above*100), type="l", xlab="depth (m)", ylab="percentage")
lines(DEM_hist$mids, (1-DEM_hist$percent_above)*100)
abline(v=-50)
abline(v=-80)

sum(mich_hist$counts[mich_hist$mids>-80&mich_hist$mids<0]) / sum(mich_hist$counts[mich_hist$mids>-50&mich_hist$mids<0])

sum(DEM_hist$counts[DEM_hist$mids>-80&DEM_hist$mids<0]) /
sum(DEM_hist$counts[DEM_hist$mids>-50&DEM_hist$mids<0])

```

# Extract volume of each bays

Total volume of the lake is $26 km^3$ (Data given in Howe et al 2006).  
Total surface of the lake is $1227 km^2$. Now I'll try counting the pixel per bays.

```{r extract volume of each bays}
l_area = 1227

DEM_missisquoi

# Missisquoi
e_missisquoi <- as(extent(-73.3, -73.0, 44.97, 45.2), 'SpatialPolygons')
crs(e_missisquoi) <- "+proj=longlat +datum=WGS84 +no_defs"
DEM_missisquoi <- crop(DEM, e_missisquoi)

plot(DEM_missisquoi)

ncell(DEM_missisquoi) / ncell(DEM) * l_area

# South lake
e_slake <- as(extent(-73.5, -73.0, 43.2, 44.2), 'SpatialPolygons')
crs(e_slake) <- "+proj=longlat +datum=WGS84 +no_defs"
DEM_slake <- crop(DEM, e_slake)
plot(DEM_slake)
ncell(DEM_slake) / ncell(DEM) * l_area

# Main lake
e_mlake <- as(extent(-73.5, -73.0, 44.2, 44.6), 'SpatialPolygons')
crs(e_mlake) <- "+proj=longlat +datum=WGS84 +no_defs"
DEM_mlake <- crop(DEM, e_mlake)
plot(DEM_mlake)
ncell(DEM_mlake) / ncell(DEM) * l_area

```

# Contour lines

```{r}
xyz <- read.delim("/Users/Rosalie/Desktop/Travail/QGIS/LakeChamplainBathy/VT_Lake_Champlain_Bathymetry.txt")
xyz$Z <- xyz$DEPTH_FT/3.281
xyz <- xyz[order(xyz$X, decreasing = F),]
contourLines(x = xyz$X, y = xyz$Y, z = xyz$Z, nlevels = 10)
#not working: error message reads Error in contourLines(x = xyz$X, y = xyz$Y, z = xyz$Z, nlevels = 10) :   increasing 'x' and 'y' values expected

# https://stackoverflow.com/questions/8406507/how-can-i-calculate-the-area-within-a-contour-in-r
library(sos)
findFn("area polygon compute")
library(splancs)
with(clines[[9]],areapl(cbind(x,y)))
with(DEM,areapl(cbind(x,y)))
areapl(poly = cbind(x= c(1:10,10:1), y=c(rep(1,10), rep(2,10))))
# Area enclosed by contour lines https://www.mathworks.com/matlabcentral/answers/132216-how-do-i-calculate-area-enclosed-by-contour-lines

# [x, y, z] = peaks; 
# [C, h] = contour(x, y, z, 20);
# ch = get(h,'children');
# area = zeros(numel(ch),1);
# for i = 1:numel(ch)
#     x = get(ch(i),'xdata');
#     y = get(ch(i),'ydata');
#     [x,y] = poly2cw(x,y);
#     area(i) = polyarea(x,y);
# end

```

