---
title: "Plankton data"
author: "Rosalie Bruel"
date: "02/05/2019"
output: 
  html_document: 
    df_print: paged
    fig_caption: yes
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
  df_print: paged
fontsize: 11pt
---

_Last updated: `r Sys.Date()`_

# Read data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(plyr)
library(dplyr)
library(gtable) # to get two plots aligned ggplot
library(grid) # low-level grid functions are required
library(plotly)
library(reshape2)


getpath4data <- function() {
  if(Sys.getenv("USER")=="Rosalie") return("/Volumes/-/Script R/LCM_GitHub_Data_LCM/")
  if(Sys.getenv("USER")=="alexnaccarato") return("~/Desktop/Food-Web 2018-2019/LCM_GitHub_Data/")
  if(Sys.getenv("USER")!="Rosalie"|Sys.getenv("USER")!="alexnaccarato") stop("You need to get the data.")
}

```

```{r load datasets, echo=FALSE, message=FALSE, warning=FALSE}
plktn_pre2010  <- read.delim(paste0(getpath4data(),"LCM_bio_PeteStangel/Plankton data pre 2010.txt"))
plktn_post2010 <- read.delim(paste0(getpath4data(),"LCM_bio_PeteStangel/Plankton data after 2010.txt"))
sites_raw      <- read.delim(paste0(getpath4data(),"LCM_bio_PeteStangel/Plankton data stations.txt"))
sites          <- sites_raw[sites_raw$LocationID %in% as.factor(plktn_pre2010$LocationID),]

plktn <- rbind(plktn_pre2010,plktn_post2010)

plktn$VisitDate <- as.Date(plktn$PlanktonData.VisitDate, format="%d-%b-%y")
head(plktn)

names(plktn)
plktn$PlanktonType[plktn$PlanktonType=="phyto"] <- "Phyto"
summary(plktn$PlanktonType)

summary(plktn$ResultType[plktn$PlanktonType=="Phyto"])
summary(plktn$ResultType[plktn$PlanktonType=="Zoo"])
summary(plktn$SampleType[plktn$PlanktonType=="Zoo"])

```


# Size to weight conversion

We use Mc Cauley 1994 work (Chapter 7 of ...).
Average sizes were obtained by browsing the web for pictures of individuals.


```{r calculate weight from size, echo=FALSE, message=FALSE, warning=FALSE}
# Size data, from McCauley 1984 (Chapter 7)
size_zoop      <- read.delim(paste0(getpath4data(),"plankton/Zoop_size.txt"))
# Calculate average weight to get biomass
size_zoop$weight <- exp(size_zoop$a +  size_zoop$b*log(size_zoop$Average_Size))
size_zoop
```


# Separate zoo and phyto

```{r message=FALSE, warning=FALSE}
# Reshape the dataframe with the type of plankton+name as column name
plktn$Type_SpeciesID <- paste(plktn$PlanktonType, plktn$SpeciesID, sep="_")
plktn2 <- dcast(plktn,  VisitDate + LocationID ~ Type_SpeciesID,value.var = "Result",fun.aggregate = sum, na.rm = TRUE )
head( plktn2)
dim(plktn2)
names(plktn2)

phyto <- plktn2[,c(1, 2, grep("Phyto_", names(plktn2)))]
zoo   <- plktn2[,c(1, 2, grep("Zoo_",   names(plktn2)))]

dim(zoo)
dim(phyto)

names(zoo)
rowSums(phyto[,-c(1:2)])

# Often, when we get 0 for zoo, we get value for phyto, and vice versa.
# no overlap of sampling? Are zoo and phyto samples taken on different days?
# There are only 69 days for which we got values both for zoo and phyto.
summary(which(rowSums(zoo[,-c(1,2)])!=0) %in% which(rowSums(phyto[,-c(1,2)])!=0))
summary(which(rowSums(phyto[,-c(1,2)])!=0) %in% which(rowSums(zoo[,-c(1,2)])!=0))


```

# Work on zooplankton

```{r group zoop species}
# Group the data by at least genus
names(zoo)
whatsleft <- 
  names(zoo[,-c(grep("aphni",     names(zoo)),
              grep("osmin",       names(zoo)),
              grep("epto",        names(zoo)),
              grep("otifer|Anuraeopsis|Ascomorpha|Asplanchna|Brachi|Collotheca|Conochilus|Cupelopagis|Euchlanis|Filinia|Kellicottia|Keratella|Lecane|Monostyla|Nothalca|Ploesoma|Polyarthra|Synchaeta|Trichocerca",    names(zoo)),
              grep("iaphanoso",               names(zoo)),
              grep("yclop",                   names(zoo)),
              grep("alanoid|Limnocalanus",    names(zoo)),
              grep("opepo|Epischura|Harpacticoid",names(zoo)))])
zoo$BOSMINA      <- rowSums(zoo[,grep("osmin",names(zoo))])     # small grazer
zoo$DAPHNIA      <- rowSums(zoo[,grep("aphni",names(zoo))])     # medium grazer
zoo$DIAPHANOSOMA <- rowSums(zoo[,grep("iaphanoso",names(zoo))]) # medium grazer
zoo$LEPTO        <- rowSums(zoo[,grep("epto", names(zoo))])      # predator
zoo$CYCLOP       <- rowSums(zoo[,grep("yclop",names(zoo))])     # copepod
zoo$CALANOID     <- rowSums(zoo[,grep("alanoid|Limnocalanus",
                                              names(zoo))])   # copepod
zoo$ROTIFER      <- rowSums(zoo[,grep("otifer|Anuraeopsis|Ascomorpha|Asplanchna|Brachi|Collotheca|Conochilus|Cupelopagis|Euchlanis|Filinia|Kellicottia|Keratella|Lecane|Monostyla|Nothalca|Ploesoma|Polyarthra|Synchaeta|Trichocerca",names(zoo))])    # small grazer
zoo$COPEPOD      <- rowSums(zoo[,grep("opepo|Epischura|Harpacticoid",names(zoo))])    # large grazer


# Group by very large groups (i.e. grazers, predators)
zoo$PREDATOR    <- rowSums(zoo[,grep("Polyphemus|LEPTO|Holopedium",names(zoo))])   
zoo$GRAZER_sm   <- rowSums(zoo[,grep("BOSMINA|Alon|Chyd|ROTIF|leurox|ladocer",names(zoo))])  
zoo$GRAZER_lr   <- rowSums(zoo[,grep("DAPHNIA|Sida|DIAPHA|CYCLOP|CALANO|opep|Eurycercus|aupl",names(zoo))])  
zoo$PARASITIC   <- zoo[,grep("Ergasilus",names(zoo))]
  
whatsleft[-c(grep("Polyphemus|LEPTO|Holopedium",whatsleft),
            grep("BOSMINA|Sida|Alon|Chyd|ROTIF|leurox|ladocer",whatsleft),
            grep("DAPHNIA|DIAPHA|CYCLOP|CALANO|COPEPOD|Eurycercus|aupl",whatsleft),
            grep("Ergasilus",whatsleft))]

plot(zoo$VisitDate,zoo$LEPTO, pch=20 )
points(zoo$VisitDate,zoo$DAPHNIA, pch=20, col=adjustcolor("blue", alpha.f = .3) )
points(zoo$VisitDate,zoo$BOSMINA, pch=20, col=adjustcolor("red", alpha.f = .3) )
points(zoo$VisitDate,zoo$PARASITIC, pch=20, col=adjustcolor("green", alpha.f = .3) )

plot(zoo$VisitDate,zoo$PARASITIC, pch=20)

summary(as.factor(plktn2$LocationID))

library(ggplot2)
ggplot(data = zoo, aes(x = VisitDate, y = PREDATOR)) + geom_point() +
  facet_wrap(~as.factor(zoo$LocationID)) + ylim(0,20000)

ggplot(data = zoo, aes(x = VisitDate, y = GRAZER_lr)) + geom_point() +
  facet_wrap(~as.factor(zoo$LocationID)) + ylim(0,100000)

ggplot(data = zoo, aes(x = VisitDate, y = GRAZER_sm)) + geom_point() +
  facet_wrap(~as.factor(zoo$LocationID)) + ylim(0,250000)

sites


# Add explanatory variables to my dataframe
zoo$LocationID2  <- sites[match(zoo$LocationID, sites$LocationID), "StationID"]
zoo$Year         <- as.numeric(format(as.Date(zoo$VisitDate, "%Y-%m-%d"),"%Y"))
zoo$Month        <- as.numeric(format(as.Date(zoo$VisitDate, "%Y-%m-%d"),"%m"))
zoo$yday         <- as.numeric(format(as.Date(zoo$VisitDate, "%Y-%m-%d"),"%j"))

# Model species
library(mgcv)
mod1 <- gamm(GRAZER_sm ~ s(Year, by=as.factor(LocationID2))+s(yday, by=as.factor(LocationID2)), data=zoo)
summary(mod1$gam)
plot(mod1$gam)

#mod2 <- gamm(GRAZER_sm ~ s(GRAZER_lr, by=as.factor(LocationID2)), data=zoo)
mod2 <- gamm(GRAZER_sm ~ s(GRAZER_lr), data=zoo)
mod2 <- gamm(GRAZER_sm ~ s(PREDATOR), data=zoo)
summary(mod2$gam)
plot(mod2$gam)

# Focus on main lake
zoo19 <- zoo[zoo$LocationID2==19,c("VisitDate", "LocationID2", "Year", "Month", "yday",
                                   "PREDATOR", "GRAZER_sm", "GRAZER_lr", "PARASITIC")]

ggplot(data = zoo19, aes(x = VisitDate, y = PARASITIC)) + geom_point()

mod2 <- lm(PREDATOR ~ Year*yday, data=zoo19)
summary(mod2)

plot(zoo19$VisitDate[zoo19$PREDATOR != 0],zoo19$PREDATOR[zoo19$PREDATOR != 0], pch=20, col=adjustcolor("black", alpha.f = .3) )
points(zoo19$VisitDate[zoo19$GRAZER_lr != 0],zoo19$GRAZER_lr[zoo19$GRAZER_lr != 0], pch=20, col=adjustcolor("blue", alpha.f = .3) )
points(zoo19$VisitDate[zoo19$GRAZER_sm != 0],zoo19$GRAZER_sm[zoo19$GRAZER_sm != 0], pch=20, col=adjustcolor("red", alpha.f = .3) )
points(zoo19$VisitDate[zoo19$PARASITIC != 0],zoo19$PARASITIC[zoo19$PARASITIC != 0], pch=20, col=adjustcolor("green", alpha.f = .3) )


#head(plktn)
#head(zoo19)

zoo19 <- zoo19[rowSums(zoo19[,c("PREDATOR", "GRAZER_sm", "GRAZER_lr", "PARASITIC")])>0,]
zoo19_month <- aggregate(zoo19[,c("PREDATOR", "GRAZER_sm", "GRAZER_lr", "PARASITIC")], list(format(zoo19$VisitDate, "%Y-%m")), mean)
#print(zoo19_month)
zoo19_month$Year   <- as.numeric(substr(zoo19_month$Group.1,1,4))
zoo19_month$Month  <- as.numeric(substr(zoo19_month$Group.1,6,7))
zoo19_month_Predator <- dcast(zoo19_month, Year ~ Month, value.var = "PREDATOR", fun.aggregate = mean)
zoo19_month_Grazer_sm <- dcast(zoo19_month, Year ~ Month, value.var = "GRAZER_sm", fun.aggregate = mean)
zoo19_month_Grazer_lr <- dcast(zoo19_month, Year ~ Month, value.var = "GRAZER_lr", fun.aggregate = mean)

```

# Phytoplankton
Investigate a bit more phytoplankton data. There are `r nrow(phyto)` observations for `r ncol(phyto)-2` groups/species.

```{r message=FALSE, warning=FALSE, include=FALSE}
dim(phyto)
countphyto <- apply(phyto[,-c(1,2)],2,function(x) length(which(x>0)))
length(countphyto[countphyto>0])
```

Out of these `r ncol(phyto)-2` groups, there are in fact only `r length(countphyto[countphyto>0])` with at least 1 observation (`r length(countphyto[countphyto>1])` with at least 2 observations, `r length(countphyto[countphyto>2])` with at least 3 observations, `r length(countphyto[countphyto>3])` with at least 4 observations).


`r ggplotly(qplot(countphyto, geom="histogram", xlab="Number of observations", ylab="Number of groups with a given number of observation", bins=max(countphyto)-min(countphyto)+1, main="Zoom in to see the details"))`

The groups with the most observations (>100) are: <br>
`r paste("- ",gsub("Phyto_","",names(countphyto[countphyto>100])), collapse="\n")`

If we decide to match a mass to only these groups, how many cell count would we lose?
```{r echo=FALSE}
phyto_summ <- data.frame("Name"= gsub("Phyto_","",colnames(phyto)[-c(1,2)]),
                        "Frequency_presence" = countphyto,
                        "Total_count" = colSums(phyto[,-c(1,2)]), row.names = NULL)

ggplotly(ggplot(phyto_summ, aes(x=Frequency_presence, y=Total_count, label=Name)) + geom_point())

#write.table(phyto_summ, "/Volumes/-/Script R/rosaliebruelweb/Blog_articles/Data/DEC_Champlain_phyto_summ.txt", col.names = T, row.names = F)
```

There are still some species we'd be losing if we chose this solution. Let's first try to match the name that are very similar.  
Trying out fuzzy matching (machine learning technique) with the <a href="https://www.google.com/search?client=safari&rls=en&ei=PWovXdr5IoXm_QaW57xg&q=fuzzy+matching+r&oq=fuzzy+matching+r&gs_l=psy-ab.3..0i203l5j0i22i10i30j0i22i30l4.15089.18111..18474...0.0..0.96.1227.16....2..0....1..gws-wiz.......0i71j35i39j0j35i39i19j0i67j0i20i263.ZVtwrl7UL6Y"> _agrep_ </a> function.

* use of _ignore.case_ to avoid problems with upper/lower case letters <br>
* use of _max.distance_ input of how many differences we allow - if we put a larger number, could be convenient to match genus? It can be an integrer, a fraction of the pattern length (maybe the best option for us? put something like 0.2?), or a list
  + argument _all_: maximal (overall) distance. If _all_ is missing, it is set to 10%, the other components default to all. The component names can be abbreviated.<br>
  + argument _insertions_: maximum number/fraction of insertions <br>
  + argument _deletions_: maximum number/fraction of deletions <br>
  + argument _substitutions_: maximum number/fraction of substitutions <br>


```{r fuzzy matching phyto name}
phyto_summ$Name_matched_num <- rep(NA,nrow(phyto_summ))
phyto_summ$Name_matched <- rep(NA,nrow(phyto_summ))
phyto_summ$Number_matched <- rep(NA,nrow(phyto_summ))

initial_vec <- seq_along(phyto_summ$Name)
edit_vec    <- initial_vec
for(i in initial_vec) {
  fuzzy_match <- 
    agrep(phyto_summ$Name[i],phyto_summ$Name,
          max = 3, ignore.case = TRUE)
  fuzzy_match <- fuzzy_match[-which(fuzzy_match==i)]
  if(length(fuzzy_match)>0) {
    phyto_summ$Name_matched_num[i] <- paste(fuzzy_match,collapse="-")
    phyto_summ$Name_matched[i] <- paste(phyto_summ$Name[fuzzy_match],collapse=", ")
    phyto_summ$Number_matched[i] <- length(fuzzy_match)     
  }
}
View(phyto_summ)
```

