---
title: "Rainbow Smelt in Lake Champlain"
author: "Rosalie Bruel and Alex Naccarato with data from Ellen Marsden"
date: "6/10/2019"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

I'm calling the libraries that will be useful for this analysis as well as cleaning up and getting familiar with the format of the data. The data was provided by Ellen Marsden. It contains 24,982 observations of smelt including important variables such as their length, weight, year class, condition, the year, and the basin it was caught in. 

```{r setup, message = FALSE}
library(ggplot2)
library(dplyr)
library(FSA)
library(FSAdata)
library(knitr)
library(leaflet)
library(janitor)
library(rgdal)
library(raster)
library(rnaturalearth)
library(RColorBrewer)

getpath4data <- function() {
  if(Sys.getenv("USER")=="Rosalie") return("/Volumes/-/Script R/LCM_GitHub_Data_LCM/")
  if(Sys.getenv("USER")=="alexnaccarato") return("~/Desktop/Food-Web 2018-2019/LCM_GitHub_Data/")
  if(Sys.getenv("USER")!="Rosalie"|Sys.getenv("USER")!="alexnaccarato") stop("You need to get the data.")
}

smelt <- read.csv(file = paste0(getpath4data(),"data_from_Ellen/MasterFileSmeltBio1984-2015-2-24-16.csv"))
smelt <- smelt[smelt$Species == "Rainbow Smelt",]
smelt$Station <- gsub("-"," ",smelt$Station)
head(smelt)

smelt_coord <- read.csv(file = paste0(getpath4data(), "smelt_coord.csv"))

xIcon <- makeIcon(
  iconUrl = "https://cdn4.iconfinder.com/data/icons/defaulticon/icons/png/256x256/cancel.png",
  iconWidth = 20, iconHeight = 20)
```

# Length-weight relationship of rainbow smelt

This is the length-weight relationship of rainbow smelt in Lake Champlain. It's consistent with plots of length-weight relationships of other fish. There are a few outliers - fish that were very overweight compared to the other smelt of similar lengths. The second plot in the output includes a trendline created from the generalized additive models method. The third plot includes a polynomial trend line.

```{r length weight relationship, results = FALSE, warning = FALSE}
# Scatterplot of weight-length relationship without a line of best fit
plot(y = smelt$Weight, x = smelt$Length, pch=20)

# Weight-length relationship of rainbow smelt with the R-generated line of best fit 
ggplot(aes(x=Length, y=Weight), data=smelt )+
  geom_point()+
  #geom_line() +
  geom_smooth(method="gam", formula = y ~ s(x))

# Weight-length relationship of rainbow smelt with a polynomial line of best fit, rather than R-generated line of best fit
ggplot(aes(x=Length, y=Weight), data=smelt )+
  geom_point() +
  stat_smooth(method = 'lm', formula = y ~ poly(x,2), se= FALSE)
```

## Getting familiar with the stations

Here, you can see rough coordinates of the stations sampled for rainbow smelt throughout the years. For ambiguous locations such as Main Lake, Main Lake North, Main Lake South, etc., we just coordinates roughly in the center of the area. You can click on the icons to see the stations specific coordinates. 

```{r, message = FALSE}
# Creation of the map with each station. Includes name of station, latitude, and longitude
coord_map <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(icon = xIcon, lng = smelt_coord$Longitude, lat = smelt_coord$Latitude,
             popup = paste("<b>Station:</b>", smelt_coord$Station, "<br>",
                           "<b>Latitude:</b>", smelt_coord$Latitude, "<br>",
                           "<b>Longitude:</b>", smelt_coord$Longitude))

coord_map
```

## Weight-length relationship of rainbow smelt by lake station

Malletts Bay includes the most outliers of any station. The smelt in the Main Lake have a particularly odd linear relationship between their length and weight from the appearance of the graph. Barber Point and the Northeast Arm have some of the largest smelt of all stations. 

```{r length weight by station, warning = FALSE}
# Grid of plots which display the weight-length relationship of rainbow smelt over time at each station
ggplot(aes(y=Weight, x=Length), data=smelt )+
  geom_point()+
  #geom_line() +
  geom_smooth(method="gam", formula = y ~ s(x)) +
  facet_wrap(~Station)
```

## Condition of rainbow smelt over time by lake station 

It appears that the condition of rainbow smelt has stayed realtively constant over time in Lake Champlain. The trendlines show an average condition of roughly 0.5 to 0.7 over time.

```{r condition by station, warning = FALSE}
# Grid of plots which display the condition of rainbow smelt over time at each station
ggplot(aes(y=condition, x=Year), data=smelt )+
  ylim(0,6)+
  geom_point()+
  #geom_line() +
  geom_smooth(method="gam", formula = y ~ s(x)) +
  facet_wrap(~Station)
```

## Abundance of species in Burlington Bay (1999-2000)

Found in the Rube Lab newsletter from early 2000's
```{r fbb, results = FALSE, warning = FALSE}
fbb <- read.delim(paste0(getpath4data(),"/data_from_Ellen/Fish_Burlington_Bay_1999_2000.txt"))
class(fbb$rel_ab_Burl_Bay_1999_2000)
fbb$abundance=gsub("<","",fbb$rel_ab_Burl_Bay_1999_2000)
fbb$abundance <- as.numeric(paste(fbb$abundance))
fbb$abundance
dim(fbb)

library(rfishbase)
dietfbb <- diet(species_list = fbb$species)
dim(dietfbb)
fbb$species[!fbb$species %in% unique(dietfbb$Species)] # all species got diet assigned
fbb$species[fbb$species %in% unique(dietfbb$Species)]
fbb$Troph<-rep(NA,nrow(fbb))
fbb$seTroph <- fbb$Troph
for (i in 1:length(unique(dietfbb$Species))) {
  fbb$Troph[i] <- mean(dietfbb$Troph[dietfbb$Species==unique(dietfbb$Species)[i]],na.rm=T)
  fbb$seTroph[i] <- mean(dietfbb$seTroph[dietfbb$Species==unique(dietfbb$Species)[i]],na.rm=T)
}
```

## Rainbow smelt per year class in Lake Champlain

The age distribution of rainbow smelt caught is skewed right. The majority of fish are 1 or 2 years old. 

```{r catch per year, warning = FALSE, message = FALSE}
# Data frame which includes the number of fish obtained per year class
summ_smelt_ageclass <- as.data.frame(summary(as.factor(smelt$Age)))
summ_smelt_ageclass <- cbind(rownames(summ_smelt_ageclass), summ_smelt_ageclass)
rownames(summ_smelt_ageclass) <- c()
colnames(summ_smelt_ageclass) <- c("year_class", "num_fish")
summ_smelt_ageclass

# Bar plot which displays the number of fish obtained per year class
ggplot(data = summ_smelt_ageclass[3:12,], aes(x = year_class, y = num_fish)) + 
  geom_bar(stat = "summary") + 
  labs(x = "Year Class", y = "Number of Fish")
```

## Sampling effort of rainbow smelt in Lake Champlain 

For 1986, 1988, and 1989, there were no data. Therefore, the number of stations and number of fish sampled for those years are equal to 0. 2002, 2003, and 2012 had particularly large sample sizes of smelt in comparison to the other years. 

In addition, the table of 0's and 1's indicates whether or not a station was sampled for a particular year. 

```{r sample effort, warning = FALSE, message = FALSE, results = "asis"}
# Data frame which includes the number of fish sampled per year
summ_smelt_effort1 <- as.data.frame(summary(as.factor(smelt$Year)))
summ_smelt_effort1 <- cbind(rownames(summ_smelt_effort1), summ_smelt_effort1)
rownames(summ_smelt_effort1) <- c()
colnames(summ_smelt_effort1) <- c("Year", "num_fish")

# Bar plot which displays the number of fish sampled per year
ggplot(data = summ_smelt_effort1, mapping = aes(x = Year, y = num_fish)) + 
  geom_bar(stat = "summary") + 
  labs(x = "Year", y = "Number of Fish Sampled") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Data frame which includes the years, list of the stations sampled during that year, and number of stations sampled that year
summ_smelt_effort2 <- smelt %>% 
  group_by(Year) %>% 
  summarise(list_stations = list(unique(Station)), num_stations = length(list_stations))

# Bar plot showing the number of stations sampled per year
ggplot(data = summ_smelt_effort2, mapping = aes(x = as.factor(Year), y = num_stations)) +
  geom_bar(stat = "summary") +
  labs(x = "Year", y = "Number of Stations Sampled") + 
  scale_y_continuous(breaks = c(seq(1, 10, 1))) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Table which displays which stations were sampled when
station_which_sampled <- as.data.frame(with(smelt, tapply(rep(1,nrow(smelt))
  ,list("Year#"=Year,"station"=Station), sum, na.rm=T)))
station_which_sampled[is.na(station_which_sampled)] <- 0
station_which_sampled[station_which_sampled > 0] <- 1

# Put in red the sites that were sampled to make it easier to see which sites were sampled when
library(huxtable)
ht <- as_hux(station_which_sampled)
ht <- set_background_color(ht, where(ht == 1), "pink")
ht <- huxtable::add_colnames(ht, colnames = colnames(station_which_sampled))
ht <- huxtable::add_rownames(ht, rownames = rownames(station_which_sampled))
col_width(ht) <- 0.5
wrap(ht) <- TRUE
ht

# Table which displays the number of fish sampled each year at each station 
station_number_sampled <- as.data.frame(with(smelt, tapply(rep(1,nrow(smelt))
  ,list("Year#"=Year,"station"=Station), sum, na.rm=T)))
station_number_sampled[is.na(station_number_sampled)] <- 0
kable(station_number_sampled)

# Table which displays the total number of fish sampled at each station across all years
station_total_sampled <- station_number_sampled %>% summarise_all(list(sum))
kable(station_total_sampled)
```



## Exploring growth parameters

### Weight-length relationship

Using Rosalie's code and Derek Ogle's tutorial.

```{r, message = FALSE, warning = FALSE}
# Two plots here--the first is the weight-length relationship of rainbow smelt without a log-transformation. The second plot is a log-transformed plot of this relationship with a line of best fit
par(mfrow=c(1,2))
plot(smelt$Weight ~ smelt$Length ,xlab="Total Length (mm)", ylab = "Weight (g)", main = "", pch=20)

smelt$logW <- log(smelt$Weight)
smelt$logL <- log(smelt$Length)

lm1 <- lm(logW~logL,data = smelt)
fitPlot(lm1,xlab="Log Total Length (mm)", ylab = "Log Weight (g)", main="")

```

The relationship is not linear without log-transforming the data. Here, we have the coefficients of the log-transformed data. 

```{r, message = FALSE, warning = FALSE}
summary(lm1)
```

### Test whether rainbow smelt exhibit isometric or allometric growth

A test of whether the fish in a population exhibit isometric growth or not can be obtained by noting that b is the estimated slope from fitting the transformed length-weight model. The slope is generically labeled with β such that the test for allometry can be translated into the following statistical hypotheses:
* H0: β=3 ⇒ H0 :"Isometricgrowth"
* HA: β≠3 ⇒ HA :"Allometricgrowth"
(All taken from Derek Ogle tutorial. Go back there for more details).

A test, and confidence interval for b, of whether rainbow smelt from Lake Champlain exhibited allometric growth or not is constructed with:

```{r, message = FALSE, warning = FALSE}
hoCoef(lm1,2,3)
confint(lm1)
```

These results show that LT exhibit allometric growth (p = 1.3e-93) with an exponent parameter (b) between 2.87 and 2.89, with 95% confidence.

### Prediction on original scale
Again, from Derek Ogle tutorial: "Predictions of the mean value of the response variable given a value of the explanatory variable can be made with predict(). In the length-weight regression, the value predicted is the mean log of weight. Most often, of course, the researcher is interested in predicting the mean weight on the original scale. An intuitive and common notion is that the log result can simply be back-transformed to the original scale by exponentiation. However, back-transforming the mean value on the log scale in this manner underestimates the mean value on the original scale. This observation stems from the fact that the back-transformed mean value from the log scale is equal to the geometric mean of the values on the original scale. The geometric mean is always less than the arithmetic mean and, thus, the back-transformed mean always underestimates the arithmetic mean from the original scale."

We want to extract the sigma, and then get the correction factor:

```{r, message = FALSE, warning = FALSE}
syx <- summary(lm1)$sigma
(cf <- exp((syx^2)/2))
```

(1) Predict log weight of a rainbow smelt of size 130 mm
(2) Biased prediction on original scale
(3) Corrected prediction on original scale

```{r, message = FALSE, warning = FALSE}
(pred.log <- predict(lm1,data.frame(logL=log(130)),interval="c")) ##(1)
(bias.pred.orig <- exp(pred.log)) ##(2)
(pred.orig <- cf*bias.pred.orig) ##(3)
```

### Comparison of weight-length relationship {.tabset}

#### Across years

Include year as a factor.  

```{r, message = FALSE, warning = FALSE}
smelt$fyear <- factor(smelt$Year)
lm2 <- lm(logW~logL*fyear, data = smelt)
```

The analysis of variable table is constructed by submitting the saved lm object to anova() as such: 

```{r, message = FALSE, warning = FALSE}
anova(lm2)
```

These results indicate that the interaction terms are significant (p = 2.2e-16). There is evidence to conclude that there is a difference in slopes in the length-weight relationship between years. The p-value for the indicator variable suggests that there is a difference in intercepts between the three years (p = 2.2e-16).

Plots and confidence intervals should be constructed for the model with the interaction term, as it was significant. The confidence intervals, constructed with:

```{r, message = FALSE, warning = FALSE}
confint(lm2)
par(mfrow=c(1,1))
fitPlot(lm2,xlab="Log Length (mm)",ylab="Log Weight (g)", legend = "topleft",main = "", col = adjustcolor(c("red","blue","grey"), alpha.f = .5))
```


#### Across stations

Include stations as a factor.  

```{r, message = FALSE, warning = FALSE}
smelt$fstations <- factor(smelt$Station)
lm3 <- lm(logW~logL*fstations, data = smelt)
``` 

The analysis of variable table is constructed by submitting the saved lm object to anova() as such:

```{r, message = FALSE, warning = FALSE}
anova(lm3)
```

These results indicate that the interaction terms is significant (p = 2.2e-16). There is evidence to conclude that there is difference in slopes in the length-weight relationship between years. The p-value for the indicator variable suggests that there is a difference in intercepts between the three sites (p = 2.2e-16). 

Plots and confidence intervals should be constructed for the model with the interaction term, as it was significant. The confidence intervals, constructed with:
```{r, message = FALSE, warning = FALSE}
confint(lm3)
par(mfrow=c(1,1))
fitPlot(lm3, xlab="Log Length (mm)",ylab="Log Weight (g)", legend = "topleft",main = "", col = adjustcolor(c("red","blue","grey"), alpha.f = .5))
```

## Conclusion: a and b parameters
In <a href=http://www.dnr.state.mi.us/publications/pdfs/ifr/manual/smii%20chapter17.pdf> Schneider et al 2010 report for Michigan Department of Natural Resources</a>, growth parameter are reported for rainbow smelt:    
  *  a = -5.12117 <br/>
  *  b =  2.96408 <br/>
We found (see results for lm1):  <br/>
  *  a = -5.05001<br/>
  *  b =  2.88230<br/>
There are some discrepancies between the two, but they're fairly close.  <br/>

To convert them to the initial equation W = aL<sup>b</sup>:  
```{r, message = FALSE, warning = FALSE}
(a = exp(lm1$coefficients[1]))
(b = lm1$coefficients[2])
```

Therefore, the intial equation is W = 0.000009405 x L<sup>2.88230</sup>

***

# Weight-length relationships for specific groups of rainbow smelt

## Age-0 (for Lake Champlain fish.xlsx)

The calculations for a and b for age-0 and age-1 rainbow smelt are for the purpose of transcribing rainbow smelt density to biomass values in Lake Champlain fish.xlsx in our Dropbox!

n = 89

Because most rainbow smelt of age-0 or with a year class of "YOY" don't have a weight, we will calculate an average weight for them based on the average length of rainbow smelt of age-0 (n = 89). We will use the weight-length equation calculated earlier for all rainbow smelt in Lake Champlain. Again, the equation from earlier was: <br/><br/> 
W = 0.000009405 x L<sup>2.88230</sup> 
<br/><br/>
Therefore, if we use the average length of 41 mm for the 89 fish with a classification of "YOY", our equation is: <br/><br/> 
W = 0.000009405 x 41<sup>2.88230</sup> 

```{r}
mean(smelt$Length[smelt$YearClass == "YOY"], na.rm = TRUE)
length(smelt$Length[smelt$YearClass == "YOY"])
(weight_age0 <- 0.000009405 * (41^2.8830))
```

The average weight of age-0 smelt in Lake Champlain is around 0.42 grams. 

## Age-1 (for Lake Champlain fish.xlsx)

n = 9,754

Using data from 9,754 rainbow smelt of age-1, we can find an average weight.

```{r}
mean(smelt$Weight[smelt$Age == 1], na.rm = TRUE)
mean(smelt$Length[smelt$Age == 1], na.rm = TRUE)
length(smelt$Weight[smelt$Age == 1])
```

The average weight of age-0 rainbow smelt in Lake Champlain is around 8.59 grams. Their average length is 114 mm.

## <4 years old

n = 10,987

```{r}
lm_4years <- lm(logW~logL,data = smelt[as.numeric(smelt$Age) <= 4,])
fitPlot(lm_4years, xlab = "Log Total Length (mm)", ylab = "Log Weight (g)", main = "")
summary(lm_4years)
```

a = -5.19480 <br/>
b =  2.96750 <br/>

W = 0.0000064 x L<sup>2.96750</sup>

## <5 years old

n = 19,412

```{r}
lm_5years <- lm(logW~logL,data = smelt[as.numeric(smelt$Age) <= 5,])
fitPlot(lm_5years, xlab = "Log Total Length (mm)", ylab = "Log Weight (g)", main = "")
summary(lm_5years)
```

a = -5.12426 <br/>
b =  2.93121 <br/>

W = 0.0000075 x L<sup>2.93121</sup>

## <6 years old

n = 23,296

```{r}
lm_6years <- lm(logW~logL,data = smelt[as.numeric(smelt$Age) <= 6,])
fitPlot(lm_6years, xlab = "Log Total Length (mm)", ylab = "Log Weight (g)", main = "")
summary(lm_6years)
```

a = -5.14460 <br/>
b =  2.90205 <br/>

W = 0.0000086 x L<sup>2.90205</sup>

## At or below 200 mm

n = 24,816

```{r}
lm_below200mm <- lm(logW~logL,data = smelt[as.numeric(smelt$Length) <= 200,])
fitPlot(lm_below200mm, xlab = "Log Total Length (mm)", ylab = "Log Weight (g)", main = "")
summary(lm_below200mm)
```

a = -5.02208 <br/>
b =  2.88082 <br/>

W = 0.0000095 x L<sup>2.880817</sup>

## Above 200 mm

n = 180

```{r}
lm_above200mm <- lm(logW~logL,data = smelt[as.numeric(smelt$Length) > 200,])
fitPlot(lm_above200mm, xlab = "Log Total Length (mm)", ylab = "Log Weight (g)", main = "")
summary(lm_above200mm)
```

a = -5.3648 <br/>
b =  3.0303 <br/>

W = 0.0000043 x L<sup>3.0303</sup>

## Summary of a and b

```{r}
# Table which includes the age/size group and their respective a and b parameters
age_group <- c("All", "<4 years", "<5 years", "<6 years", "<=200 mm", ">200 mm")
a <- c(-5.05001, -5.19480, -5.12426, -5.14460, -5.02208, -5.3648)
b <- c(2.88230, 2.96750, 2.93121, 2.90205, 2.88082, 3.0303)

a_and_b_summary <- data.frame(age_group, a, b)
kable(a_and_b_summary)
```

