---
title: "Stocking data VTFWS and NYDEC"
author: "Rosalie Bruel"
date: "15/05/2019"
output: 
  html_document: 
    df_print: paged
    fig_caption: yes
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
  df_print: paged
fontsize: 11pt
---

_Last updte: `r Sys.Date()`_

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, include=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(stringr)
library("sf")
library(rgdal) # to read bathymetry
library("rnaturalearth")
library("rnaturalearthdata")
library("ggspatial")
#load spatial packages
library(raster)
library(rgeos)
library(gtable) # to get two plots aligned ggplot
library(grid) # low-level grid functions are required
library(scatterpie) # for pie chart on map
library(lubridate) # to standardize date-time with different format

getpath4data <- function() {
  if(Sys.getenv("USER")=="Rosalie") return("/Volumes/-/Script R/LCM_GitHub_Data_LCM/")
  if(Sys.getenv("USER")=="alexnaccarato") return("~/Desktop/Food-Web 2018-2019/LCM_GitHub_Data/")
  if(Sys.getenv("USER")!="Rosalie"|Sys.getenv("USER")!="alexnaccarato") stop("You need to get the data.")
}

```


The VT Fish and Wildlife Department states that 'stocking fish helps to restore populations of native fish, improve fisheries in Vermont, and provide better fishing opportunities'.

# Accessing data for each agency {.tabset}

## VTFWD

### Read data and basic stats

Stocking data were obtained from the <a href= https://vtfishandwildlife.com/fishvt>VTFWD website</a> by clicking on the link 'Trout Stocking Report'.

```{r include=FALSE}
stkvt <- read.delim(paste0(getpath4data(),'Stocking/VTFWD_Stocking_2009-2019.txt'))
nrow(stkvt)
length(stkvt$ACTUAL.NUMBER.STOCKED[!is.na(stkvt$ACTUAL.NUMBER.STOCKED)])
names(stkvt)

stkvt$DATE.STOCKED <- parse_date_time(x = stkvt$DATE.STOCKED,orders = c("m/d/y", "m/d/Y"))
```

Over the `r min(stkvt$Year)`-`r max(stkvt$Year)` period, `r nrow(stkvt)` stocking events were scheduled, and `r length(stkvt$ACTUAL.NUMBER.STOCKED[!is.na(stkvt$ACTUAL.NUMBER.STOCKED)])` stocking events took place.

```{r echo=FALSE, message=FALSE, warning=FALSE}

summ_stkvt <- stkvt %>% group_by(as.factor(Year)) %>%
  summarize(mean=mean(ACTUAL.NUMBER.STOCKED, na.rm=T),
            sum=sum(ACTUAL.NUMBER.STOCKED, na.rm=T),
            scheduled=length(ACTUAL.NUMBER.STOCKED),
            went_through=length(ACTUAL.NUMBER.STOCKED[!is.na(ACTUAL.NUMBER.STOCKED)]),
            percent_done=went_through/scheduled)
summ_stkvt <- as.data.frame(summ_stkvt)

p1 <- ggplot(data=summ_stkvt, aes(x=summ_stkvt[,1],y=as.numeric(summ_stkvt[,2]))) +
  geom_bar(stat="identity") + xlab("Year") + ylab("mean # stocked") 

p2 <- ggplot(data=summ_stkvt, aes(x=summ_stkvt[,1],y=as.numeric(summ_stkvt[,3]))) +
  geom_bar(stat="identity") + xlab("Year") + ylab("sum # stocked") 

## convert plots to gtable objects
g1 <- ggplotGrob(p1)
g2 <- ggplotGrob(p2)
g <- rbind(g1, g2, size="first") # stack the two plots
g$widths <- unit.pmax(g1$widths, g2$widths) # use the largest widths
# center the legend vertically
g$layout[grepl("guide", g$layout$name),c("t","b")] <- c(1,nrow(g))
grid.newpage()
grid.draw(g)

colnames(summ_stkvt) <- c("Year","Mean #fish stocked", "Total #fish stocked", "Schedule events", "Stocking that went through", "% stocking done")
summ_stkvt

```

In 2019, there are still many stocking events that were not carried out. We're removing this year from the analysis.

```{r remove 2019 from the dataframe}
stkvt <- stkvt[stkvt$Year!=2019,]
```

### Species stocked

What are the species stocked by VTFWD? We're only looking at the effective stocking now.

```{r echo=FALSE, message=FALSE, warning=FALSE}

sp_stkvt <- stkvt %>% group_by(as.factor(SPECIES)) %>%
  summarize(mean=mean(ACTUAL.NUMBER.STOCKED, na.rm=T),
            sum=sum(ACTUAL.NUMBER.STOCKED, na.rm=T),
            scheduled=length(ACTUAL.NUMBER.STOCKED),
            went_through=length(ACTUAL.NUMBER.STOCKED[!is.na(ACTUAL.NUMBER.STOCKED)]),
            percent_done=went_through/scheduled)

ggplot(data=stkvt, aes(x=SPECIES,y=ACTUAL.NUMBER.STOCKED, fill=as.factor(Year))) +
  geom_bar(stat="identity") + xlab("Species") + ylab("total # stocked") +
  labs(fill="Year") + ggtitle("All Vermont")

ggplot(data=stkvt[grep("Champ",stkvt$WATER),], aes(x=SPECIES,y=ACTUAL.NUMBER.STOCKED, fill=as.factor(Year))) +
  geom_bar(stat="identity") + xlab("Species") + ylab("total # stocked") +
  labs(fill="Year") + ggtitle("Lake Champlain waters")


```


```{r message=FALSE, warning=FALSE, include=FALSE}
bathy<-readOGR(paste0(getpath4data(),"GIS/LakeChamplain_Shapefile/LakeChamplain.shp"))
lcgroupmap=c("lake",rep("island", 84562))
world <- ne_countries(scale = "medium", returnclass = "sf")

class(bathy)
crs(bathy)
crs(world)
bathy <- spTransform(bathy,crs(world))
crs(bathy)

# get states too
library(USAboundaries)
library(USAboundariesData)
states <- us_states(resolution = "high")

```

```{r include=FALSE}
# 1 - Load gps location for different cities
# Then convert them to correct projection
gps1 <- read.delim(paste0(getpath4data(),"Stocking/GPS_coordinates_stocking_locations_counties.txt"))
gps1$County <- gsub("\xca","",gps1$County)
gps1$State <- gsub("\xca","",gps1$State)
gps1_df <- gps1
coordinates(gps1)<-~Longitude+Latitude
proj4string(gps1)<-CRS("+proj=longlat +datum=WGS84")
gps1 <- spTransform(gps1,crs(world))

gps2 <- read.delim(paste0(getpath4data(),"Stocking/GPS_coordinates_stocking_locations_LC.txt"))
gps2_df <- gps2
gps2_df$Sites <- as.character(gps2_df$Sites)
coordinates(gps2)<-~Longitude+Latitude
proj4string(gps2)<-CRS("+proj=longlat +datum=WGS84")
gps2 <- spTransform(gps2,crs(world))
gps2_df$Longitude <- gps2$Longitude
gps2_df$Latitude <- gps2$Latitude


# 2 - Match gps to initial data
stkvt$lon <- rep(NA, nrow(stkvt))
stkvt$lat <- rep(NA, nrow(stkvt))
for (i in 1:nrow(stkvt)) {
  if(i==1) {n1=0;n2=0;n3=0;which_double <- NULL; number_not_matched=0;number_not_matched2=0;which_not_matched <- NULL}
  if(stkvt$COUNTY[i] %in% gps1_df$County) {
    if(length(gps1_df$Longitude[gps1_df$County==stkvt$COUNTY[i]])==1) {
      stkvt$lon[i] <- gps1_df$Longitude[gps1_df$County==stkvt$COUNTY[i]]
      stkvt$lat[i] <- gps1_df$Latitude[gps1_df$County==stkvt$COUNTY[i]]
      n1 <- n1+1 # for info message
    } else {
      n2 <- n2+1 # for info message
      which_not_matched <- c(which_not_matched, i)
      if(!is.na(stkvt$ACTUAL.NUMBER.STOCKED[i])) number_not_matched <- number_not_matched + stkvt$ACTUAL.NUMBER.STOCKED[i]
      if(length(grep(stkvt$WATER[i], which_double))==0) which_double <- paste(which_double,"       - ", stkvt$WATER[i], "\n")
    }
    
  } else {
      n3 <- n3+1 # for info message
            if(!is.na(stkvt$ACTUAL.NUMBER.STOCKED[i])) number_not_matched2 <- number_not_matched2 + stkvt$ACTUAL.NUMBER.STOCKED[i]
      }
  # Info message
  if(i==nrow(stkvt)) message(paste0(" ✓ Found GPS coordinates match for ", n1," of the stocking events","\n ✕ More than 1 match for ", n2, " of the stocking events.\n  It corresponds to ",number_not_matched," fish (out of ",sum(stkvt$ACTUAL.NUMBER.STOCKED,na.rm=T),").\n   Sites are:\n",which_double," ✕ No GPS coordinates match were found for ", n3, " of the stocking events.\n  It corresponds to ",number_not_matched2," fish (out of ",sum(stkvt$ACTUAL.NUMBER.STOCKED,na.rm=T),")." ))
}

# 3 - create a format that scatterpie will be able to handle
# It will look like:
#    ~Lon   ~Lat    ~species1   ~species2    ~species3
#      xx     xx        xxx           xxx         xxx
summ_stkvt_map <- stkvt
summ_stkvt_map$flonlat <- as.factor(paste(summ_stkvt_map$lon, summ_stkvt_map$lat, sep="___"))
summ_stkvt_map <- with(summ_stkvt_map, tapply(ACTUAL.NUMBER.STOCKED,list("Lon Lat"=flonlat, "Species"=SPECIES), sum, na.rm=T))
summ_stkvt_map <- as.data.frame(summ_stkvt_map)
summ_stkvt_map$total <- rowSums(summ_stkvt_map,na.rm=T)
summ_stkvt_map$flonlat <- rownames(summ_stkvt_map)
rownames(summ_stkvt_map) <- 1:nrow(summ_stkvt_map)
summ_stkvt_map <- cbind(summ_stkvt_map[,1:(ncol(summ_stkvt_map)-1)],data.frame(do.call('rbind', strsplit(as.character(summ_stkvt_map$flonlat), '___', fixed=TRUE))))
colnames(summ_stkvt_map) <- str_replace(colnames(summ_stkvt_map), "X1", "lon")
colnames(summ_stkvt_map) <- str_replace(colnames(summ_stkvt_map), "X2", "lat")
summ_stkvt_map$lon <- as.numeric(paste(summ_stkvt_map$lon))
summ_stkvt_map$lat <- as.numeric(paste(summ_stkvt_map$lat))

summ_stkvt_map <- summ_stkvt_map[!is.na(summ_stkvt_map$lon),]
summ_stkvt_map[is.na(summ_stkvt_map)] <- 0
```

Here are the results from matching GPS coordinates to initial dataframe:

```{r echo=FALSE}
message(paste0(" ✓ Found GPS coordinates match for ", n1," of the stocking events","\n ✕ More than 1 match for ", n2, " of the stocking events.\n  It corresponds to ",number_not_matched," fish (out of ",sum(stkvt$ACTUAL.NUMBER.STOCKED,na.rm=T),").\n   Sites are:\n",which_double," ✕ No GPS coordinates match were found for ", n3, " of the stocking events.\n  It corresponds to ",number_not_matched2," fish (out of ",sum(stkvt$ACTUAL.NUMBER.STOCKED,na.rm=T),")." ))

```



```{r echo=FALSE, message=FALSE, warning=FALSE}

# library(scatterpie) for pie charts on map

p <- ggplot() +  
  geom_sf(data = world)  + 
  geom_sf(data = states) +
  coord_sf(xlim = c(-77, -70), ylim = c(40, 46))  +
  theme_set(theme_bw()) +
  xlab("Longitude") + ylab("Latitude")
# p <- p +
#   geom_polygon(data=bathy, aes(x=long, y=lat, group=group, fill=lcgroupmap), show.legend = FALSE) +  scale_fill_manual(values = c("lightgrey","darkgrey"))
p <- p +
    geom_scatterpie(data = summ_stkvt_map, 
                    aes(x=lon, y=lat, r = sqrt(total+4000)/4000),
                    cols = c("BKT", "BNT", "LAS", "LAT", "RBT", "STT", "WAE"),  alpha = 0.5) +
    scale_fill_manual(
        breaks = c("BKT", "BNT", "LAS", "LAT", "RBT", "STT", "WAE"),
        labels = c("BKT", "BNT", "LAS", "LAT", "RBT", "STT", "WAE"),
        values = c("BKT" = "orange",
                   "BNT" = "red",
                   "LAS" = "darkred",
                   "LAT" = "yellow",
                   "RBT" = "grey",
                   "STT" = "black",
                   "WAE" = "cyan")#,
        # labels = c("Brook Trout", "Brown Trout", "Landlocked Atlantic Salmon", "Lake Trout", "Rainbow Trout", "Steelhead Trout")
        )

p

```


## NYDEC

### Read data and basic stats

Stocking data were obtained from the <a href= https://data.ny.gov/Recreation/Fish-Stocking-Lists-Actual-Beginning-2011/9hpx-asd8>NYDEC website</a>.

```{r include=FALSE}
stkny <- read.delim(paste0(getpath4data(),"Stocking/NYDEC_Stocking_2011-2019.txt"))
nrow(stkny)
names(stkny)

#QAQC some data
stkny$Town[stkny$Town=="plattsburgh"] <- "Plattsburgh"

```

Over the `r min(stkny$Year)`-`r max(stkny$Year)` period, `r nrow(stkny)` stocking events took place.

```{r echo=FALSE, message=FALSE, warning=FALSE}

summ_stkny <- stkny %>% group_by(as.factor(Year)) %>%
  summarize(mean=mean(Number, na.rm=T),
            sum=sum(Number, na.rm=T),
            went_through=length(Number))
summ_stkny <- as.data.frame(summ_stkny)

p1 <- ggplot(data=summ_stkny, aes(x=summ_stkny[,1],y=as.numeric(summ_stkny[,2]))) +
  geom_bar(stat="identity") + xlab("Year") + ylab("mean # stocked") 

p2 <- ggplot(data=summ_stkny, aes(x=summ_stkny[,1],y=as.numeric(summ_stkny[,3]))) +
  geom_bar(stat="identity") + xlab("Year") + ylab("sum # stocked") 

## convert plots to gtable objects
g1 <- ggplotGrob(p1)
g2 <- ggplotGrob(p2)
g <- rbind(g1, g2, size="first") # stack the two plots
g$widths <- unit.pmax(g1$widths, g2$widths) # use the largest widths
# center the legend vertically
g$layout[grepl("guide", g$layout$name),c("t","b")] <- c(1,nrow(g))
grid.newpage()
grid.draw(g)

colnames(summ_stkny) <- c("Year","Mean #fish stocked", "Total #fish stocked", "Stocking events")
summ_stkny

```


### Species stocked

What are the species stocked by VTFWD? We're only looking at the effective stocking now.

```{r echo=FALSE, message=FALSE, warning=FALSE}

sp_stkny <- stkny %>% group_by(as.factor(Species)) %>%
  summarize(mean=mean(Number, na.rm=T),
            sum=sum(Number, na.rm=T),
            went_through=length(Number))

ggplot(data=stkny, aes(x=Species,y=Number, fill=as.factor(Year))) +
  geom_bar(stat="identity") + xlab("Species") + ylab("total # stocked") +
  labs(fill="Year") + ggtitle("All NY")

ggplot(data=stkny[grep("Champ",stkny$Waterbody),], aes(x=Species,y=Number, fill=as.factor(Year))) +
  geom_bar(stat="identity") + xlab("Species") + ylab("total # stocked") +
  labs(fill="Year") + ggtitle("Lake Champlain waters")


```


# Focus on Lake Champlain

## VT data

Grand Isle gathers the most stocking events over the years, but it's in Colchester that most fish were stocked.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Vermont data
stkvtlc <- stkvt[grep("Champ",stkvt$WATER),]
#head(stkvtlc)

stkvtsites <- stkvtlc %>% group_by(TOWN) %>%
  summarize(meanFishStocked=mean(ACTUAL.NUMBER.STOCKED, na.rm=T),
            sumFishStocked=sum(ACTUAL.NUMBER.STOCKED, na.rm=T),
            sumFishStockedPerYear=sum(ACTUAL.NUMBER.STOCKED, na.rm=T)/10,
            NumberStockingEvents=length(ACTUAL.NUMBER.STOCKED))

stkvtsites[order(stkvtsites$sumFishStocked, decreasing = T),]

#levels(stkvtlc$COUNTY)

```

## NY data:

```{r echo=FALSE}
# NY data
stknylc <- stkny[grep("hampl",stkny$Waterbody),]

stknylcsites <- stknylc %>% group_by(Town) %>%
  summarize(meanFishStocked=mean(Number, na.rm=T),
            sumFishStocked=sum(Number, na.rm=T),
            sumFishStockedPerYear=sum(Number, na.rm=T)/7,
            NumberStockingEvents=length(Number))

stknylcsites[order(stknylcsites$sumFishStocked, decreasing = T),]

```



## Mapping


```{r echo=FALSE, message=FALSE, warning=FALSE}
summ_sp_yr_vt <- as.data.frame(with(stkvtlc, tapply(ACTUAL.NUMBER.STOCKED
  ,list("species"=SPECIES,"Year#"=Year), sum, na.rm=T)))
summ_sp_yr_vt <- as.data.frame(t(summ_sp_yr_vt[rowSums(summ_sp_yr_vt, na.rm=T)>0,]))
summ_sp_yr_vt$Year <- rownames(summ_sp_yr_vt)
summ_sp_yr_vt$State <- "VT"

summ_sp_yr_ny <- as.data.frame(with(stknylc, tapply(Number
  ,list("species"=Species,"Year#"=Year), sum, na.rm=T)))
summ_sp_yr_ny <- as.data.frame(t(summ_sp_yr_ny[rowSums(summ_sp_yr_ny, na.rm=T)>0,]))

colnames(summ_sp_yr_ny) <- str_replace(colnames(summ_sp_yr_ny), "Brown Trout", "BNT")
colnames(summ_sp_yr_ny) <- str_replace(colnames(summ_sp_yr_ny), "Landlocked Salmon", "LAS")
colnames(summ_sp_yr_ny) <- str_replace(colnames(summ_sp_yr_ny), "Lake Trout", "LAT")
colnames(summ_sp_yr_ny) <- str_replace(colnames(summ_sp_yr_ny), "Round Whitefish", "RWI")

summ_sp_yr_ny$Year <- rownames(summ_sp_yr_ny)
summ_sp_yr_ny$State <- "NY"


custom_rbind = function(x1,x2){
  c1 = setdiff(colnames(x1),colnames(x2))
  c2 = setdiff(colnames(x2),colnames(x1))
  for(c in c2){##Adding missing columns from 2 in 1
    x1[[c]]=NA
  }
  for(c in c1){##Similiarly ading missing from 1 in 2
    x2[[c]]=NA
  }
  x2 = x2[colnames(x1)]
  rbind(x1,x2)
}

summ_sp_yr <- Reduce(custom_rbind,list(summ_sp_yr_vt,summ_sp_yr_ny))  
rownames(summ_sp_yr) <- paste(summ_sp_yr$Year, summ_sp_yr$State, sep="_")

```

The dataframe below gives the average lentgh of fish released this year. The average length of fish changed over the season, so I balanced the length by the proportion released on that month vs. fish from the same specie release over the year.

```{r details on species stocked by NY, echo=FALSE, message=FALSE, warning=FALSE}
stknylc$Size_pondere <- rep(NA, nrow(stknylc))
for (i in 1:nrow(stknylc)) {
  stknylc$Size_pondere[i] <- stknylc$Size_mm[i]*stknylc$Number[i]/sum(stknylc$Number[stknylc$Year==stknylc$Year[i] & stknylc$Species==stknylc$Species[i]])
}

summ_sp_yr_ny <- as.data.frame(with(stknylc, tapply(Size_pondere
  ,list("species"=Species,"Year#"=Year), sum, na.rm=T)))
summ_sp_yr_ny <- as.data.frame(t(summ_sp_yr_ny[rowSums(summ_sp_yr_ny, na.rm=T)>0,]))

summ_sp_yr_ny

```

## Map 

```{r message=FALSE, warning=FALSE, include=FALSE}
# 1 - gps data were loaded previously

# 2a - Match gps to initial data for VT
stkvtlc$TOWN[stkvtlc$TOWN=="panton"] = "Panton"
stkvtlc$lon <- rep(NA, nrow(stkvtlc))
stkvtlc$lat <- rep(NA, nrow(stkvtlc))
for (i in 1:nrow(stkvtlc)) {
  if(i==1) {n1=0;n2=0;n3=0;which_double <- NULL;which_abs <- NULL; number_not_matched=0;number_not_matched2=0;which_not_matched <- NULL}
  if(stkvtlc$TOWN[i] %in% gps2_df$Sites) {
    if(length(gps2_df$Longitude[gps2_df$Sites==stkvtlc$TOWN[i]])==1) {
      stkvtlc$lon[i] <- gps2_df$Longitude[gps2_df$Sites==stkvtlc$TOWN[i]]
      stkvtlc$lat[i] <- gps2_df$Latitude[gps2_df$Sites==stkvtlc$TOWN[i]]
      n1 <- n1+1 # for info message
    } else {
      n2 <- n2+1 # for info message
      which_not_matched <- c(which_not_matched, i)
      if(!is.na(stkvtlc$ACTUAL.NUMBER.STOCKED[i])) number_not_matched <- number_not_matched + stkvtlc$ACTUAL.NUMBER.STOCKED[i]
      if(length(grep(stkvtlc$WATER[i], which_double))==0) which_double <- paste(which_double,"       - ", stkvtlc$WATER[i], "\n")
    }
    
  } else {
      n3 <- n3+1 # for info message
      if(!is.na(stkvt$ACTUAL.NUMBER.STOCKED[i])) number_not_matched2 <- number_not_matched2 + stkvt$ACTUAL.NUMBER.STOCKED[i]
      if(length(grep(stkvtlc$WATER[i], which_double))==0) which_abs <- paste(which_abs,"       - ", stkvtlc$WATER[i], "\n")
      }
  # Info message
  if(i==nrow(stkvtlc)) message(paste0(" ✓ Found GPS coordinates match for ", n1," of the stocking events","\n ✕ More than 1 match for ", n2, " of the stocking events.\n  It corresponds to ",number_not_matched," fish (out of ",sum(stkvt$ACTUAL.NUMBER.STOCKED,na.rm=T),").\n   Sites are:\n",which_double," ✕ No GPS coordinates match were found for ", n3, " of the stocking events.\n  It corresponds to ",number_not_matched2," fish (out of ",sum(stkvt$ACTUAL.NUMBER.STOCKED,na.rm=T),")..\n  Sites are:\n",which_abs ))
}

```

Results GPS matching for VT:

```{r echo=FALSE, warning=FALSE}
message(paste0(" ✓ Found GPS coordinates match for ", n1," of the stocking events","\n ✕ More than 1 match for ", n2, " of the stocking events.\n  It corresponds to ",number_not_matched," fish (out of ",sum(stkvt$ACTUAL.NUMBER.STOCKED,na.rm=T),").\n   Sites are:\n",which_double," ✕ No GPS coordinates match were found for ", n3, " of the stocking events.\n  It corresponds to ",number_not_matched2," fish (out of ",sum(stkvt$ACTUAL.NUMBER.STOCKED,na.rm=T),")..\n  Sites are:\n",which_abs ))
```


```{r message=FALSE, warning=FALSE, include=FALSE}

# 3 - create a format that scatterpie will be able to handle
# It will look like:
#    ~Lon   ~Lat    ~species1   ~species2    ~species3
#      xx     xx        xxx           xxx         xxx
summ_stkvtlc_map <- stkvtlc
summ_stkvtlc_map$flonlat <- as.factor(paste(summ_stkvtlc_map$lon, summ_stkvtlc_map$lat, sep="___"))
summ_stkvtlc_map <- with(summ_stkvtlc_map, tapply(ACTUAL.NUMBER.STOCKED,list("Lon Lat"=flonlat, "Species"=SPECIES), sum, na.rm=T))
summ_stkvtlc_map <- as.data.frame(summ_stkvtlc_map)
summ_stkvtlc_map$total <- rowSums(summ_stkvtlc_map,na.rm=T)
summ_stkvtlc_map$flonlat <- rownames(summ_stkvtlc_map)
rownames(summ_stkvtlc_map) <- 1:nrow(summ_stkvtlc_map)
summ_stkvtlc_map <- cbind(summ_stkvtlc_map[,1:(ncol(summ_stkvtlc_map)-1)],data.frame(do.call('rbind', strsplit(as.character(summ_stkvtlc_map$flonlat), '___', fixed=TRUE))))
colnames(summ_stkvtlc_map) <- str_replace(colnames(summ_stkvtlc_map), "X1", "lon")
colnames(summ_stkvtlc_map) <- str_replace(colnames(summ_stkvtlc_map), "X2", "lat")
summ_stkvtlc_map$lon <- as.numeric(paste(summ_stkvtlc_map$lon))
summ_stkvtlc_map$lat <- as.numeric(paste(summ_stkvtlc_map$lat))

summ_stkvtlc_map <- summ_stkvtlc_map[!is.na(summ_stkvtlc_map$lon),]
summ_stkvtlc_map[is.na(summ_stkvtlc_map)] <- 0


```


```{r include=FALSE}

# 2b - Match gps to initial data for NY
stknylc$Town <- as.character(stknylc$Town)
stknylc$Town[stknylc$Town=="Plattsburgh"] <- "Plattsburgh NY"
stknylc$Town[stknylc$Town=="Port Kent"]   <- "Port Kent NY"
stknylc$Town[stknylc$Town=="Moriah"]      <- "Moriah NY"
stknylc$Town[stknylc$Town=="Chesterfield, Essex"]      <- "Chesterfield, Essex NY"
stknylc$Town[stknylc$Town=="Essex"]      <- "Chesterfield, Essex NY"
stknylc$lon <- rep(NA, nrow(stknylc))
stknylc$lat <- rep(NA, nrow(stknylc))
for (i in 1:nrow(stknylc)) {
  if(i==1) {n1=0;n2=0;n3=0;which_double <- NULL;which_abs<-NULL; number_not_matched=0;number_not_matched2=0;which_not_matched <- NULL}
  if(stknylc$Town[i] %in% gps2_df$Sites) {
    if(length(gps2_df$Longitude[gps2_df$Sites==stknylc$Town[i]])==1) {
      stknylc$lon[i] <- gps2_df$Longitude[gps2_df$Sites==stknylc$Town[i]]
      stknylc$lat[i] <- gps2_df$Latitude[gps2_df$Sites==stknylc$Town[i]]
      n1 <- n1+1 # for info message
    } else {
      n2 <- n2+1 # for info message
      which_not_matched <- c(which_not_matched, i)
      if(!is.na(stknylc$Number[i])) number_not_matched <- number_not_matched + stknylc$Number[i]
      if(length(grep(stknylc$Town[i], which_double))==0) which_double <- paste(which_double,"       - ", stknylc$Town[i], "\n")
    }
    
  } else {
    n3 <- n3+1 # for info message
    if(!is.na(stknylc$Number[i])) number_not_matched2 <- number_not_matched2 + stknylc$Number[i]
    if(length(grep(stknylc$Town[i], which_abs))==0) which_abs <- paste(which_abs,"       - ", stknylc$Town[i], "\n")
  }
  # Info message
  if(i==nrow(stknylc)) message(paste0(" ✓ Found GPS coordinates match for ", n1," of the stocking events","\n ✕ More than 1 match for ", n2, " of the stocking events.\n  It corresponds to ",number_not_matched," fish (out of ",sum(stknylc$Number,na.rm=T),").\n  Sites are:\n",which_double," ✕ No GPS coordinates match were found for ", n3, " of the stocking events.\n  It corresponds to ",number_not_matched2," fish (out of ",sum(stknylc$Number,na.rm=T),").\n  Sites are:\n",which_abs ))
}

```

Results GPS matching for NY:

```{r echo=FALSE}
message(paste0(" ✓ Found GPS coordinates match for ", n1," of the stocking events","\n ✕ More than 1 match for ", n2, " of the stocking events.\n  It corresponds to ",number_not_matched," fish (out of ",sum(stknylc$Number,na.rm=T),").\n  Sites are:\n",which_double," ✕ No GPS coordinates match were found for ", n3, " of the stocking events.\n  It corresponds to ",number_not_matched2," fish (out of ",sum(stknylc$Number,na.rm=T),").\n  Sites are:\n",which_abs ))
```


```{r message=FALSE, warning=FALSE, include=FALSE}

# 3 - create a format that scatterpie will be able to handle
# It will look like:
#    ~Lon   ~Lat    ~species1   ~species2    ~species3
#      xx     xx        xxx           xxx         xxx
summ_stknylc_map <- stknylc
summ_stknylc_map$flonlat <- as.factor(paste(summ_stknylc_map$lon, summ_stknylc_map$lat, sep="___"))
summ_stknylc_map <- with(summ_stknylc_map, tapply(Number,list("Lon Lat"=flonlat, "Species"=Species), sum, na.rm=T))
summ_stknylc_map <- as.data.frame(summ_stknylc_map)
summ_stknylc_map$total <- rowSums(summ_stknylc_map,na.rm=T)
summ_stknylc_map$flonlat <- rownames(summ_stknylc_map)
rownames(summ_stknylc_map) <- 1:nrow(summ_stknylc_map)
summ_stknylc_map <- cbind(summ_stknylc_map[,1:(ncol(summ_stknylc_map)-1)],data.frame(do.call('rbind', strsplit(as.character(summ_stknylc_map$flonlat), '___', fixed=TRUE))))
colnames(summ_stknylc_map) <- str_replace(colnames(summ_stknylc_map), "X1", "lon")
colnames(summ_stknylc_map) <- str_replace(colnames(summ_stknylc_map), "X2", "lat")
summ_stknylc_map$lon <- as.numeric(paste(summ_stknylc_map$lon))
summ_stknylc_map$lat <- as.numeric(paste(summ_stknylc_map$lat))
# remove rows with no GPS coordinates 
summ_stknylc_map <- summ_stknylc_map[!is.na(summ_stknylc_map$lat),]
# remove species column with no count
summ_stknylc_map[is.na(summ_stknylc_map)] <- 0

summ_stknylc_map <-summ_stknylc_map[,abs(colSums(summ_stknylc_map))>0]

colnames(summ_stknylc_map) <- str_replace(colnames(summ_stknylc_map), "Brown Trout", "BNT")
colnames(summ_stknylc_map) <- str_replace(colnames(summ_stknylc_map), "Landlocked Salmon", "LAS")
colnames(summ_stknylc_map) <- str_replace(colnames(summ_stknylc_map), "Lake Trout", "LAT")

```


```{r match ny and vt gps datasets, include=FALSE}
# write function to add extra column when the name doesn't exist yet
matchColName <- function(df_origin,df_toedit,replace=NA) {
  for (i in 1:ncol(df_origin)) {
    if (i==1) n1=0
    if (!colnames(df_origin)[i] %in% colnames(df_toedit)) {
      df_toedit$newcol <- rep(replace, nrow(df_toedit))
      colnames(df_toedit) <- str_replace(colnames(df_toedit), "newcol", colnames(df_origin)[i])
      n1=n1+1
    }
    if(i==ncol(df_origin)) message(paste0(n1," columns were added to the second data frame to match the first one columns."))
  }
  df_toedit <- return(df_toedit)
}

# apply function to 2 dataframes
summ_stkvtlc_map <- matchColName(summ_stknylc_map,summ_stkvtlc_map, replace=0)
summ_stknylc_map <- matchColName(summ_stkvtlc_map,summ_stknylc_map, replace=0)

# rbind dataframes
summ_stkvtlc_map$Agency <- "VT"
summ_stknylc_map$Agency <- "NY"
summ_stklc_map <- rbind(summ_stkvtlc_map,summ_stknylc_map[,colnames(summ_stkvtlc_map)])

# when lon or lat are very close, make only one point
summ_stklc_map <- summ_stklc_map[order(summ_stklc_map$lon),]
newdf <- summ_stklc_map
for (i in which(duplicated(summ_stklc_map$lon)&duplicated(summ_stklc_map$lat))) {
  newdf[i,] <- c(colSums(summ_stklc_map[summ_stklc_map$lat==summ_stklc_map$lat[i],!colnames(summ_stklc_map) %in% c("lon","lat","Agency")]),summ_stklc_map[i,colnames(summ_stklc_map) %in% c("lon","lat","Agency")])
  newdf$Agency[i] <- "NY/VT" 
}
summ_stklc_map <- newdf[-(which(duplicated(summ_stklc_map$lon))-1),]


```

Fish stocked on Lake Champlain watershed. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
p <- ggplot() +  
  #geom_sf(data = world) +
  geom_sf(data = states) +
  coord_sf(xlim = c(-74, -72.5), ylim = c(43.5, 45.5))  +
  theme_set(theme_bw()) +
  xlab("Longitude") + ylab("Latitude")

p <- p + geom_polygon(data=bathy, aes(x=long, y=lat, group=group, fill=lcgroupmap), show.legend = FALSE) +
  scale_fill_manual(values = c("lightgrey","darkgrey"))

p  +
  geom_scatterpie(data = summ_stklc_map, 
                    aes(x=lon, y=lat, r = sqrt(total)/12000),
                    cols = c("BKT", "BNT", "LAS", "LAT", "RBT", "STT", "WAE"),  alpha = 0.5) +
    scale_fill_manual(
        breaks = c("BKT", "BNT", "LAS", "LAT", "RBT", "STT", "WAE"),
        labels = c("BKT", "BNT", "LAS", "LAT", "RBT", "STT", "WAE"),
        values = c("BKT" = "orange",
                   "BNT" = "red",
                   "LAS" = "darkred",
                   "LAT" = "yellow",
                   "RBT" = "grey",
                   "STT" = "black",
                   "WAE" = "cyan")
    )

```

