---
title: "Zooplankton of Lake Champlain"
author: "Alex Naccarato"
date: "8/6/2019"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(plyr)
library(dplyr)
library(gtable) # to get two plots aligned ggplot
library(grid) # low-level grid functions are required
library(plotly)
library(reshape2)
library(lubridate)
library(plotly)

getpath4data <- function() {
  if(Sys.getenv("USER")=="Rosalie") return("/Volumes/-/Script R/LCM_GitHub_Data_LCM/")
  if(Sys.getenv("USER")=="alexnaccarato") return("~/Desktop/Food-Web 2018-2019/LCM_GitHub_Data/")
  if(Sys.getenv("USER")!="Rosalie"|Sys.getenv("USER")!="alexnaccarato") stop("You need to get the data.")
}

plktn_pre2010  <- read.delim(paste0(getpath4data(),"LCM_bio_PeteStangel/Plankton data pre 2010.txt"))
plktn_post2010 <- read.delim(paste0(getpath4data(),"LCM_bio_PeteStangel/Plankton data after 2010.txt"))
sites_raw      <- read.delim(paste0(getpath4data(),"LCM_bio_PeteStangel/Plankton data stations.txt"))
sites          <- sites_raw[sites_raw$LocationID %in% as.factor(plktn_pre2010$LocationID),]

plktn <- rbind(plktn_pre2010,plktn_post2010)

plktn$VisitDate <- parse_date_time(x = plktn$PlanktonData.VisitDate, orders = c("d-b-y","d/m/Y"))
head(plktn)

names(plktn)
plktn$PlanktonType[plktn$PlanktonType=="phyto"] <- "Phyto"

plktn$Type_SpeciesID <- paste(plktn$PlanktonType, plktn$SpeciesID, sep="_")
plktn2 <- dcast(plktn,  VisitDate + LocationID ~ Type_SpeciesID,value.var = "Result",fun.aggregate = sum, na.rm = TRUE )
head( plktn2)
dim(plktn2)
names(plktn2)

phyto <- plktn2[,c(1, 2, grep("Phyto_", names(plktn2)))]
zoo   <- plktn2[,c(1, 2, grep("Zoo_",   names(plktn2)))]

zoop_param <- read.delim(paste0(getpath4data(), "zoop_parameters.txt"))

whatsleft <- 
  names(zoo[,-c(grep("aphni",     names(zoo)),
              grep("osmin",       names(zoo)),
              grep("epto",        names(zoo)),
              grep("otifer|Anuraeopsis|Ascomorpha|Asplanchna|Brachi|Collotheca|Conochilus|Cupelopagis|Euchlanis|Filinia|Kellicottia|Keratella|Lecane|Monostyla|Nothalca|Ploesoma|Polyarthra|Synchaeta|Trichocerca",    names(zoo)),
              grep("iaphanoso",               names(zoo)),
              grep("yclop",                   names(zoo)),
              grep("alanoid|Limnocalanus",    names(zoo)),
              grep("opepo|Epischura|Harpacticoid",names(zoo)))])
zoo$BOSMINA      <- rowSums(zoo[,grep("osmin",names(zoo))])     # small grazer
zoo$DAPHNIA      <- rowSums(zoo[,grep("aphni",names(zoo))])     # medium grazer
zoo$DIAPHANOSOMA <- rowSums(zoo[,grep("iaphanoso",names(zoo))]) # medium grazer
zoo$LEPTO        <- rowSums(zoo[,grep("epto", names(zoo))])      # predator
zoo$CYCLOP       <- rowSums(zoo[,grep("yclop",names(zoo))])     # copepod
zoo$CALANOID     <- rowSums(zoo[,grep("alanoid|Limnocalanus",
                                              names(zoo))])   # copepod
zoo$ROTIFER      <- rowSums(zoo[,grep("otifer|Anuraeopsis|Ascomorpha|Asplanchna|Brachi|Collotheca|Conochilus|Cupelopagis|Euchlanis|Filinia|Kellicottia|Keratella|Lecane|Monostyla|Nothalca|Ploesoma|Polyarthra|Synchaeta|Trichocerca",names(zoo))])    # small grazer
zoo$COPEPOD      <- rowSums(zoo[,grep("opepo|Epischura|Harpacticoid",names(zoo))])    # large grazer


# Group by very large groups (i.e. grazers, predators)
zoo$PREDATOR    <- rowSums(zoo[,grep("Polyphemus|LEPTO|Holopedium",names(zoo))])   
zoo$GRAZER_sm   <- rowSums(zoo[,grep("BOSMINA|Alon|Chyd|ROTIF|leurox|ladocer",names(zoo))])  
zoo$GRAZER_lr   <- rowSums(zoo[,grep("DAPHNIA|Sida|DIAPHA|CYCLOP|CALANO|opep|Eurycercus|aupl",names(zoo))])  
zoo$PARASITIC   <- zoo[,grep("Ergasilus",names(zoo))]
  
whatsleft[-c(grep("Polyphemus|LEPTO|Holopedium",whatsleft),
            grep("BOSMINA|Sida|Alon|Chyd|ROTIF|leurox|ladocer",whatsleft),
            grep("DAPHNIA|DIAPHA|CYCLOP|CALANO|COPEPOD|Eurycercus|aupl",whatsleft),
            grep("Ergasilus",whatsleft))]

zoo$LocationID2  <- sites[match(zoo$LocationID, sites$LocationID), "StationID"]
zoo$Year         <- as.numeric(format(as.Date(zoo$VisitDate, "%Y-%m-%d"),"%Y"))
zoo$Month        <- as.numeric(format(as.Date(zoo$VisitDate, "%Y-%m-%d"),"%m"))
zoo$yday         <- as.numeric(format(as.Date(zoo$VisitDate, "%Y-%m-%d"),"%j"))

# Biomass values      <- old zoop counts           ug/m3    g/m3      g/km3        t/km3
zoo$BOSMINA_biom      <- round(zoo$BOSMINA      *  12.252 / 1000000 * 1000000000 / 1000000 , 2)
zoo$DAPHNIA_biom      <- round(zoo$DAPHNIA      * 195.881 / 1000000 * 1000000000 / 1000000 , 2)
zoo$DIAPHANOSOMA_biom <- round(zoo$DIAPHANOSOMA *  26.219 / 1000000 * 1000000000 / 1000000 , 2)
zoo$LEPTO_biom        <- round(zoo$LEPTO        * 163.714 / 1000000 * 1000000000 / 1000000 , 2)
zoo$CYCLOP_biom       <- round(zoo$CYCLOP       *  96.724 / 1000000 * 1000000000 / 1000000 , 2)
zoo$CALANOID_biom     <- round(zoo$CALANOID     * 229.355 / 1000000 * 1000000000 / 1000000 , 2)
zoo$ROTIFER_biom      <- round(zoo$ROTIFER      *   7.861 / 1000000 * 1000000000 / 1000000 , 2)
zoo$COPEPOD_biom      <- round(zoo$COPEPOD      *  19.126 / 1000000 * 1000000000 / 1000000 , 2)
zoo$PREDATOR_biom     <- round(zoo$PREDATOR     * 213.582 / 1000000 * 1000000000 / 1000000 , 2)
zoo$GRAZER_sm_biom    <- round(zoo$GRAZER_sm    *  21.039 / 1000000 * 1000000000 / 1000000 , 2)
zoo$GRAZER_lr_biom    <- round(zoo$GRAZER_lr    * 187.096 / 1000000 * 1000000000 / 1000000 , 2)
zoo$PARASITIC_biom    <- round(zoo$PARASITIC    *  35.039 / 1000000 * 1000000000 / 1000000 , 2)
```

## Summary values for zooplankton weights

### Mean wet weight of each large group

<span style="color:red">Specify the unit, either in the title or with a sentence here.</span>

```{r}
# Prints the average wet weight for each of the large groups 
for (i in unique(zoop_param$Large_group)) {
  print(paste0(i, ": ", round(mean(zoop_param$WW__g[zoop_param$Large_group == i], na.rm = TRUE), 3)))
}
```

### Mean wet weight of each small group

```{r}
# Prints the wet weight for each of the small groups 
for (i in unique(zoop_param$Small_group)) {
  print(paste0(i, ": ", (round(mean(zoop_param$WW__g[zoop_param$Small_group == i], na.rm = TRUE), 3))))
}
```

## Overall dataset summary

### Distribution of the average size of zooplankton

Here, we're plotting two very similar plots. The first contains the zooplankton <i> Eurycercus spp. </i> which has an average wet weight of over 3,000 micrograms. The second does not contain this same data point so the viewer can better observe the majority of the data.

<span style="color:red">Good idea to plot them with and without Eurycercus. A good way to look at the distribution would also to visualize <a href="http://www.sthda.com/english/wiki/wiki.php?title=ggplot2-density-easy-density-plot-using-ggplot2-and-r-statistical-software">dentsity plots</a>.</span>


```{r}
# Histogram of wet weight of all species (includes Eurycercus spp.)
ggplot(data = zoop_param, mapping = aes(x = WW__g)) +
  geom_histogram(bins = 30) +
  labs(x = expression(Wet~Weight~(mu*'g')), y = "Count")

zoop_param <- zoop_param[zoop_param$WW__g < 2000,]

# Histogram of wet weight all of species (does not include Eurycercus spp.)
ggplot(data = zoop_param, mapping = aes(x = WW__g)) +
  geom_histogram(bins = 30) +
  labs(x = expression(Wet~Weight~(mu*'g')), y = "Count")

zoop_param <- read.delim(paste0(getpath4data(), "zoop_parameters.txt"))
```

## Intraannual changes

<span style="color:red">Have a fixes x-axis (minimum is 5 I believe, max is 10? May-October?), it will ease comparison.</span>

```{r}
# Function that plots the counts of zooplankton by each month across all years
counts_all_month <- 
  function(year) {
    zoop_sub <- zoo[zoo$Year == year,]
    print(ggplot(data = zoop_sub, mapping = aes(x = zoop_sub$Month, y = rowSums(zoop_sub[, 84:95]))) +
      geom_bar(stat = "identity") +
      labs(y = "Total Zooplankton Count", x = "Month"))
  }

# Function that plots the biomass of zooplankton by each month across all years
biom_all_month <- 
  function(year) {
    zoop_sub <- zoo[zoo$Year == year,]
    print(ggplot(data = zoop_sub, mapping = aes(x = zoop_sub$Month, y = rowSums(zoop_sub[, 100:111]))) +
      geom_bar(stat = "identity") +
      labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month"))
  }

# Function that plots the counts of each zooplankton category by each month across all years
counts_group_month <- 
  function(group, group_names) {
    print(ggplot(data = zoo, mapping = aes(x = Month, y = group)) +
      geom_bar(stat = "identity") +
      labs(y = "Total Zooplankton Count", x = "Month"))
  }

# Function that plots the biomass of each zooplankton category by each month across all years 
biom_group_month <- 
  function(group, group_names) {
    print(ggplot(data = zoo, mapping = aes(x = Month, y = group)) +
      geom_bar(stat = "identity") +
      labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Month"))
  }
```

### Counts and biomass for each year

```{r, results = 'asis'}
# Only 1993-2013 because 1992, 2014, and 2015 show empty plots
for (i in 1993:2013) {
  cat('\n')  
  cat("####Counts for", i, "\n") 
  counts_all_month(year=i)
  cat('\n') 
  cat('\n')  
  cat("####Biomass for", i, "\n") 
  biom_all_month(year=i)
  cat('\n')
}
```

### Counts by group

```{r, results = 'asis'}
for (i in 84:95) {
  cat('\n')  
  cat("####Counts for", names(zoo[i]), " by month \n")
  counts_group_month(group = zoo[,i], group_names = names(zoo[i]))
  cat('\n') 
}
```

### Biomass by group

```{r, results = 'asis'}
for (i in 100:111) {
  cat('\n')  
  cat("####Biomass for", names(zoo[i]), " by month \n")
  biom_group_month(group = zoo[,i], group_names = names(zoo[i]))
  cat('\n') 
  }
```

## Interannual changes

### Counts and biomass over time 

<span style="color:red">Has the effort changed? Maybe it'd be usefull to add a section with some metadata at the beginning: sites that were sampled, how many times a year, etc. Also, see when they were sampled: since the abundance vary seasonally (your month plots above), maybe only compare august counts (we have august every year right?). <br>
Another idea would be to add the 'station' as factor in the plot, and plot it with plotly (see example in the LTdiet.Rmd or stocking.Rmd). If we can select/unselect stations, we could compare the abundance at a single site. <br>
If it varies, we'll have to standardize the rest of the data. </span>

```{r}
for (i in unique(zoo$Year)) {
  print(paste0(i, ": ", sum(rowSums(zoo[zoo$Year == i, 84:95])), " zooplankton"))
}

# Plots the zooplankton counts over time
ggplot(data = zoo, mapping = aes(x = zoo$Year, y = rowSums(zoo[,84:95]))) +
  geom_bar(stat = "identity") +
  labs(y = "Total Zooplankton Count")

for (i in unique(zoo$Year)) {
  print(paste0(i, ": ", sum(rowSums(zoo[zoo$Year == i, 100:111])), " t/km3"))
}

# Plots the zooplankton biomass over time
ggplot(data = zoo, mapping = aes(x = zoo$Year, y = rowSums(zoo[, 100:111]))) +
  geom_bar(stat = "identity") +
  labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Year")
```


```{r}
# Function that plots the counts of each category over time
counts_group_year <- 
  function(group, group_names) {
    print(ggplot(data = zoo, mapping = aes(x = Year, y = group)) +
      geom_bar(stat = "identity") +
      labs(y = expression(Total~Zooplankton~Biomass~(t/km^3)), x = "Year"))
  }

# Function that plots the biomass of each category over time
biom_group_year <- 
  function(group, group_names) {
    print(ggplot(data = zoo, mapping = aes(x = Year, y = group)) +
      geom_bar(stat = "identity") +
      labs(y = "Total Zooplankton Count", x = "Year"))
  }
```

### Counts by group

```{r, results = 'asis'}
for (i in 84:95) {
  cat('\n')  
  cat("####Counts for", names(zoo[i]), " by year \n") 
  counts_group_year(group = zoo[,i], group_names = names(zoo[i]))
  cat('\n')
}
```

### Biomass by group

```{r, results = 'asis'}
for (i in 100:111) {
  cat('\n')  
  cat("####Biomass for", names(zoo[i]), " by year \n") 
  biom_group_year(group = zoo[,i], group_names = names(zoo[i]))
  cat('\n')
}
```

